---
title: "Clean LENA 5 min data"
output: html_notebook
---

##Load data
```{r load_data}
library(tidyverse)
library(lubridate)
library(hms)

#daily files
#lena_daily_file_1 <- "~/Desktop/BABIES/r01_winter_2019/data/LENA Export (Daily) 20180316 154518.csv"
lena_daily_file_1 <- "~/Desktop/lena_18mo_symptoms/Data/LENA Export (Daily) 20180316 154518_ALL.csv"

#lena_daily_file_2 <- "~/Desktop/BABIES/r01_winter_2019/data/LENA Export (Daily) 20181128 153328.csv"
lena_daily_file_2 <- "~/Desktop/lena_18mo_symptoms/Data/LENA Export (Daily) 20181128 153328.csv"
lena_daily_file_3 <- "~/Desktop/lena_18mo_symptoms/Data/LENAExport_Daily_new_20190206.csv"

#5-minute files
#lena_5min_file_1 <- "~/Desktop/BABIES/r01_winter_2019/data/LENAExport_5Minute_20181220.csv"
lena_5min_file_1 <- "~/Desktop/lena_18mo_symptoms/Data/LENAExport_5Minute_20181220_cleaned.csv"
lena_5min_file_2 <- "~/Desktop/lena_18mo_symptoms/Data/LENAExport_5Minute_new_20190204.csv"


#lena_5min_files_2 <- "~/Desktop/BABIES/r01_winter_2019/data/individual_5min"
lena_5min_files_2 <- "~/Desktop/lena_18mo_symptoms/Data/individual_5min_lsk_format_cleaned"

lena_followup_long_file <- "~/Desktop/lena_18mo_symptoms/Data/lena_followup_long_20190216.csv"
lena_followup_cs_file <- "~/Desktop/lena_18mo_symptoms/Data/lena_followup_cs_20190216.csv"

lena_5min_files_df <-
  tibble(
    path = 
      list.files(
        path = lena_5min_files_2,
        all.files = TRUE,
        #pattern = "^LENA Export (5 Minute) \\d{3}_\\d{6}.csv$",
        no.. = TRUE,
        full.names = TRUE
      )
  ) %>% 
  filter(
    str_detect(path, "/.DS_Store") == FALSE
  )
```
#5 min data
#Reformat sp 5 min data
```{r}
lena_5min_sp <-
  read_csv(lena_5min_file_2) %>% 
  mutate(
    ExternalReferenceID = as.integer(ExternalReferenceID),
    Recording_DOB = as.character(Recording_DOB),
    StartTime = as.character(StartTime)
  ) %>% 
  select(
    ExternalReferenceID,
    Recording_DOB,
    StartTime,
    Duration_Secs,
    AWC_COUNT,
    CT_COUNT,
    CV_COUNT,
    Recording_Gender,
    Meaningful,
    Distant, 
    TV_Elec,
    Overlap,
    Noise,
    Silence
  ) %>% 
  mutate(
    ID = ExternalReferenceID,
    dob = Recording_DOB,
    AWC.Actual = AWC_COUNT,
    CTC.Actual = CT_COUNT,
    CVC.Actual = CV_COUNT,
    Sex = Recording_Gender,
    Duration = as.hms(Duration_Secs),
    Meaningful = as.hms(Meaningful),
    Distant = as.hms(Distant),
    TV = as.hms(TV_Elec),
    Overlap = as.hms(Overlap),
    Noise = as.hms(Noise),
    Silence = as.hms(Silence),
    Timestamp = StartTime
  ) %>% 
  select(
    -ExternalReferenceID,
    -Recording_DOB,
    -Recording_Gender,
    -CT_COUNT,
    -CV_COUNT,
    -AWC_COUNT,
    -Duration_Secs
  )
```

# Read in 5-minute data from Pro license 2 (2018 file) and SP
```{r warning=FALSE}
lena_5min <-
  read_csv(lena_5min_file_1) %>% 
  mutate(
    Birthdate = as.character(Birthdate),
    Timestamp = as.character(Timestamp),
    Meaningful = as.hms(Meaningful)
  ) %>% 
  select(
    -Type,
    -ChildKey,
    -Id,
    -Age,
    -AVA_StdScore:-AVA_AvgScore_Pct,
    -Firstname,
    ID = Lastname,
    dob = Birthdate
  ) %>% 
  bind_rows(
    lena_5min_sp
  ) %>% 
  separate(Timestamp, into = c("date_record", "time_record"), sep = " ") %>% 
  mutate(
    ID = as.numeric(ID),
    dob = parse_date_time(dob, c("ymd", "mdy")),
    date_record = parse_date_time(date_record, c("ymd", "mdy")),
    age = (dob %--% date_record) / months(1)
  ) %>% 
  select(
    -dob
  )
```
#Read in data from individual files (Pro license 1)
```{r warning=FALSE}
lena_5min2 <-
  lena_5min_files_df %>%
  mutate(
    data = map(
      path, 
      read_csv, 
      col_names = TRUE,
      col_types = 
        list(
          Firstname = col_character(),
          Birthdate = col_character(),
          Timestamp = col_character(),
          AVA_StdScore = col_character(),
          Lastname = col_character(),
          Sex = col_character(),
          ChildKey = col_character()
        )
      )
  ) %>% 
  unnest(data) %>% 
  select(-path) %>% 
  mutate(
    Birthdate = as.character(Birthdate),
    Timestamp = as.character(Timestamp)
  ) %>% 
  select(
    -Type,
    -ChildKey,
    -Id,
    -Age,
    -AVA_StdScore:-AVA_AvgScore_Pct,
    ID1 = Firstname,
    ID2 = Lastname,
    dob = Birthdate
  ) %>% 
  separate(Timestamp, into = c("date_record", "time_record"), sep = " ") %>% 
  mutate(
    ID = as.numeric(
      if_else(
        is.na(ID2),
        ID1, ID2
      )
    ),
    dob = parse_date_time(dob, c("ymd", "mdy")),
    date_record = parse_date_time(date_record, c("ymd", "mdy")),
    age = (dob %--% date_record) / months(1)
  ) %>% 
  select(
    ID,
    everything(),
    -ID1,
    -ID2,
    -dob
  )
```
#combine all data
```{r}
lena_5min <-
  lena_5min %>% 
  bind_rows(lena_5min2) %>% 
  arrange(ID)
```

## Clean and wrangle 5-minute data (generate prop and daily rate variables)
#conditions to filter on: age between 5 and 8 months at time of recording, duration of recording at least 8 hours
```{r}
lena_5min_cl <- 
  lena_5min %>% 
  group_by(ID) %>% 
  mutate(
    duration = as.numeric(
      as.duration(hms(as.numeric(Duration))),
      "hours"
    )
  ) %>% 
  group_by(ID, date_record, age) %>% 
  summarise(
    duration = sum(duration),
    awc_depend = sum(AWC.Actual >= 1),
    ctc_depend = sum(CTC.Actual >= 1),
    cvc_depend = sum(CVC.Actual >= 1),
    awc_total = sum(AWC.Actual),
    ctc_total = sum(CTC.Actual),
    cvc_total = sum(CVC.Actual)
  ) %>% 
  mutate(
    awc_daily_rate = awc_total/duration, #daily AWC rate
    ctc_daily_rate = ctc_total/duration, #daily CTC rate
    cvc_daily_rate = cvc_total/duration, #daily CVC rate
    awc_prop = (awc_depend) / ((duration * 60) / 5), #proportion of 5 minute segments with AWC
    ctc_prop = (ctc_depend) / ((duration * 60) / 5), #proportion of 5 minute segments with CTC 
    cvc_prop = (cvc_depend)/ ((duration * 60) / 5)
  ) %>% 
  filter(
    (age >= 5 & age <= 8),
    duration >= 8
  ) %>% 
  group_by(ID) %>% 
  mutate(
    day_type = wday(date_record),
    day_type = if_else(day_type == 1 | day_type == 7, "weekend", "weekday"),
    day = 1:n()
  ) %>% 
  ungroup()
```

#create wideform dataframe with 1 row per ID
```{r}
lena_5min_cl %>% 
  count(day)

lena_5min_day1_rename <-
  lena_5min_cl %>% 
  filter(day == 1) %>% 
  rename_at(
    vars(date_record:day_type),
    funs(paste0(., "_D1"))
  ) %>% 
  select(-day)

lena_5min_day2_rename <-
  lena_5min_cl %>% 
  filter(day == 2) %>% 
  rename_at(
    vars(date_record:day_type),
    funs(paste0(., "_D2"))
  ) %>% 
  select(-day)

lena_5min_cl_wf <-
  lena_5min_day1_rename %>% 
  left_join(lena_5min_day2_rename, by = "ID") %>% 
  mutate(
    #day_type_D1 = wday(date_record_D1),
    #day_type_D1 = if_else(day_type_D1 == 1 | day_type_D1 == 7, "weekend", "weekday"),
    #day_type_D2 = wday(date_record_D2),
    #day_type_D2 = if_else(day_type_D2 == 1 | day_type_D2 == 7, "weekend", "weekday")
  )

lena_5min_cl_wf %>% 
  filter(
    awc_prop_D1 > .2,
    awc_prop_D2 > .2
    ) %>% 
  ggplot(aes(awc_prop_D1, awc_prop_D2)) +
  geom_point() +
  geom_smooth(method = "lm")

glimpse(lena_5min_cl_wf)
```
##identifying weird recordings and what to do about them
```{r}
##Deleted from raw data: 1052 5/8/18 recording, 62 Day 1 on 7/11/17 (from LENAExport_5minute_20181220)
##Filter from analyses: 56 Day 1, 110 Day 2, use 106 day 1 (day 2 not typical), use 1052 day 2 (day 1 "went to zoo for 90 min" but doesn't say when), use 1045 day 2 (day 1 not whole day), 1010 use day 2 (mother worked 5 hrs day 1)

##Deleting data with crowds -cleaned from raw data (files renamed "_cleaned")
#recordings with crowds --
  #cs: 27 9/18/16 15-16:00 (5 min files), 119 4/19/18 11:30-12:30 (20181220), 135 9/13/18 13-13:30 (20181220), 38 2/10/17 16-18 (20181220), 48 3/5/17 11-13 (20181220), 61 7/18/17 13-15 (20181220), 62 7/30/17 9:30-10:30 (2018), 76 8/21/17 15-16 (2018), 2 day 1 19:30-10 (5 min files), 28 day 1 13-15 (5 min files), 64 8/8/17 11-11:30 (2018), 75 9/15/17 12-15 (2018), 87 9/11/17 14:40-16:40 (2018), 105 3/15/18 16:45-16:55 (2018), 27 10/2/16 11-13 (5 min files), 119 5/12/18 13:30-17:30 (2018), 135 9/23/18 9-10:45 (2018), 12 day 2 12:20-13:50 (5 min files), 45 4/22/17 10:30-11:30 (2018), 50 3/18/17 9:45-12:45 (not completed correctly; entirely deleted), 60 7/16/17 14:30-16:30 (2018), 72 8/6/17 13-16 (2018), 94 4/15/18 12:30-14 (2018), 106 4/1/18 14:40-18:40 (2018), 115 4/14/18 13-14:30 (2018)
  #long: 1029 4/6/18 18:40-20:40 (2018), 1021 6/25/18 18-19:30 (2018), 1078 6/15/18 17:30-19:30 (2018), 1028 4/17/18 15:30-16 (2018), 1036 3/22/18 13:15-15:15 (2018), 1045 4/22/18 10-12 (2018), 1010 2/8/18 9:15-14:15, 1028 5/6/18 15-19 (2018), 1036 5/5/18 8-9 16-17 (2018), 1045 5/19/18 10:30-12:30 (2018), 1011 5/13/18 15-16 (2018), 1019 5/20/18 16-17 (2018), 1023 5/12/18 11-13 (2018), 1060 9/29/18 17:30-18:30 (2018)


```
#count weekdays vs weekends
```{r}
lena_5min_cl_wf %>% 
  count(day_type_D1)

lena_5min_cl_wf %>% 
  count(day_type_D2)

lena_5min_cl_wf %>% 
  filter(day_type_D2 == "weekend" | day_type_D1 == "weekend")


#Conclusion: 54 with weekend recording, so need to covary for day type (can't just use weekend from everyone); also many more weekends in day 2 than day 1
```
#summarize proportion variables
```{r}
lena_5min_cl_wf %>% 
  summarise_at(
    vars(awc_prop_D1, ctc_prop_D1),
    funs(min, max, mean, sd), na.rm = TRUE
  )
```
#read in LENA follow up data
```{r}
lena_followup <-
  read_csv(lena_followup_long_file) %>% 
  rename(ID = record_id) %>% 
  mutate(ID = as.integer(ID)) %>% 
  bind_rows(
    read_csv(lena_followup_cs_file) %>% 
      rename(ID = record_id) %>% 
      mutate(ID = as.integer(ID))
  ) %>% 
  filter(
    lena_day_1_followup_questions_complete == 2 | lena_day_2_followup_questions_complete == 2
  ) %>% 
  mutate(
    lena_weekend = lena_daytype
  )

#t-test weekday vs. weekend and percent of time spent w mom
t.test(percent_mother ~ lena_weekend, data = lena_followup)
```
###merge relevant follow up qs data with cleaned & filtered lena data - **unfinished**
```{r}
##add percent_mother variable to lena data to see if difference weekday vs. weekend within the subset of recordings we've filtered, joining on ID and day_type

lena_5min_cl_filtered <-
  lena_5min_cl_filtered %>% 
  mutate(
    day_weekend_num = if_else(
      day_type == "weekend",
      1, 0
    )
  )
  

left_join(
    lena_followup
  ) %>% 
  select(
    percent_mother
  )
```

##Write file with relevant data 

```{r}
#all data
write_csv(lena_5min_cl, "~/Desktop/lena_18mo_symptoms/Data/lena_cleaned_byday_20190202.csv")
write_csv(lena_5min_cl_wf, "~/Desktop/lena_18mo_symptoms/Data/lena_cleaned_byID_20190202.csv")

#Filtered data to use for lena_18month_symptoms project 
#IDs to use day 2: 
# day 2: 110 (put device in pocket rather than chest day 1), 106

#110 day 2 isn't in dataset and 106 isn't in it at all, so taking day 1 from everyone:

lena_5min_cl_filtered <-
  lena_5min_cl %>%
  filter(
    ID != 110
  ) %>% 
  filter(
    day == 1
  )
  

  
write_csv(lena_5min_cl_filtered, "~/Desktop/lena_18mo_symptoms/Data/lena_cleaned_day1_20190202.csv")
  
```