---
title: "lena_18mo_symptoms_analyses"
author: "Fran Querdasi"
date: "3/2/2019"
output: html_document
---

```{r setup, include=FALSE}
#Parameters

##set up CRAN
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

#load libraries
library(tidyverse)
library(Hmisc)
library(car)
library(corrr)
library(corrplot)
library(lme4)
library(ggpubr)
library(modelr)

#get files

##Fran file paths
crisys_file <- "~/Desktop/lena_18mo_symptoms/Data/crisys_6mo_20190430.csv"
q_18mo_file <- "~/Desktop/lena_18mo_symptoms/Data/q_18month_scores_20190510.csv"
q_18mo_ages_file <- "~/Desktop/lena_18mo_symptoms/Data/q_18month_ages.csv"
demographics_file <- "~/Desktop/lena_18mo_symptoms/Data/demographics_20190510.csv"
lena_file <- "~/Desktop/lena_18mo_symptoms/Data/lena_cleaned_day1_20190510.csv"
lena_follow_and_data_file <- "~/Desktop/lena_18mo_symptoms/Data/lena_followup_and_data_20190510.csv"
baby_raceth_file <- "~/Desktop/lena_18mo_symptoms/Data/baby_race_ethnicity.csv"
ibq_file <- "~/Desktop/lena_18mo_symptoms/Data/ibq_scored_20190510.csv"
lscr_file <- "~/Desktop/lena_18mo_symptoms/Data/lscr_merged_scored_20190412.csv"

##Lucy file paths
# crisys_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/crisys_6mo_20190430.csv"
# q_18mo_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/q_18month_scores_20190413.csv"
# q_18mo_ages_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/q_18month_ages.csv"
# demographics_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/demographics_20190413.csv"
# lena_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/lena_cleaned_day1_20190413.csv"
# lena_follow_and_data_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/lena_followup_and_data_20190413.csv"
# baby_raceth_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/baby_race_ethnicity.csv"
# ibq_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/ibq_scored_20190413.csv"
# lscr_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/Data_20190413/lscr_merged_scored_20190412.csv"
```

# Read in and join data
```{r}
lena_18mo_sym_df <-
  read_csv(q_18mo_file) %>% 
  select(
    ID, 
    itsea_symptoms,
    itsea_competence,
    baby_age_18month
  ) %>% 
  left_join(
    read_csv(crisys_file) 
  ) %>% 
  left_join(
    read_csv(lscr_file) %>% 
      select(
        ID,
        num_stress_total,
        num_stress_preg,
        sum_sev,
        max_sev
      )
  ) %>% 
  left_join(
    read_csv(demographics_file) %>% 
      select(
        ID,
        education,
        annual_income,
        income_needs,
        male,
        mom_age_6mo_visit,
        mom_race,
        mom_ethnicity,
        child_race,
        childrace_describe,
        child_ethnicity
      )
  ) %>% 
  left_join(
    read_csv(lena_file) %>% 
      rename(
        baby_age_lena = age
      )
  ) %>% 
  left_join(
    read_csv(ibq_file) %>% 
      select(
        ID,
        SUR,
        NEG
      )
  )

lena_18mo_sym_df <-
  lena_18mo_sym_df %>% 
  filter(
    !is.na(day),
    !(ID %in% c(1019, 1020) & is.na(mom_age_6mo_visit)) #these were duplicates
    ) %>% 
  unique()  

glimpse(lena_18mo_sym_df)
  
```
#Descriptives in full dataset

Race:
1, American Indian or Alaska Native
2, Asian
3, Black or African American
4, Native Hawaiian or Other Pacific Islander
5, White
6, Other

Education:
0, No schooling completed
2, Nursery school to 8th grade
3, Some high school, no diploma
4, High school graduate, diploma or the equivalent (i.e. GED)
5, Some college credit, no degree
6, Trade/technical/vocational training
7, Associate degree
8, Bachelor's degree
9, Graduate degree
10, Other

Income:
1, $0-5,000
2, $5,001-15,000
3, $15,001-30,000
4, $30,001-60,000
5, $60,001-90,000
6, $90,001-150,000
7, Greater than $150,000
```{r}
#means and sds of continuous demographic variables
demo_summaries <-
lena_18mo_sym_df %>% 
  summarize_at(
    vars(
    baby_age_18month,
    baby_age_lena,
    duration,
    mom_age_6mo_visit
    ), 
    list(~mean, ~sd, ~min, ~max),
    na.rm = TRUE)

#counts and percentages for categorical variables (education, ethnicity, race, income)

#income
lena_18mo_sym_df %>% 
  count(annual_income) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(annual_income))

lena_18mo_sym_df %>% 
  count(income_needs < 1) %>% 
  mutate(per = n / sum(n))

#mom and child ethnicity
lena_18mo_sym_df %>% 
  count(mom_ethnicity) %>% 
  mutate(per = n / sum(n))

lena_18mo_sym_df %>% 
  count(child_ethnicity) %>% 
  mutate(per = n / sum(n))

#mom and child race
lena_18mo_sym_df %>% 
  count(mom_race) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(per))

lena_18mo_sym_df %>% 
  count(child_race) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(per))

#mom education
lena_18mo_sym_df %>% 
  count(education) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(per))
```

# Descriptives in subset with complete data

```{r}
lena_18mo_sym_df_complete <-
  lena_18mo_sym_df %>% 
  filter(
    complete.cases(lena_18mo_sym_df)
  )

#means and sds of continuous demographic variables
demo_summaries_complete <-
lena_18mo_sym_df_complete %>% 
  summarize_at(
    vars(
    baby_age_18month,
    baby_age_lena,
    duration,
    mom_age_6mo_visit
    ), 
    list(~mean, ~sd, ~min, ~max),
    na.rm = TRUE)

#counts and percentages for categorical variables (education, ethnicity, race, income)

#income
lena_18mo_sym_df_complete %>% 
  count(annual_income) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(annual_income))

lena_18mo_sym_df_complete %>% 
  count(income_needs < 1) %>% 
  mutate(per = n / sum(n))

#mom and child ethnicity
lena_18mo_sym_df_complete %>% 
  count(mom_ethnicity) %>% 
  mutate(per = n / sum(n))

lena_18mo_sym_df_complete %>% 
  count(child_ethnicity) %>% 
  mutate(per = n / sum(n))
 
#mom and child race 
lena_18mo_sym_df_complete %>% 
  count(mom_race) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(per))

lena_18mo_sym_df %>% 
  count(child_race) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(per))

#mom education
lena_18mo_sym_df_complete %>% 
  count(education) %>% 
  mutate(per = n / sum(n)) %>% 
  arrange(desc(per))
```

# Look at distributions of data
```{r}
qplot(lena_18mo_sym_df$income_needs, 
      geom = "histogram",
      binwidth = .1)

qplot(lena_18mo_sym_df$education, 
      geom = "histogram",
      binwidth = 1)

qplot(lena_18mo_sym_df$baby_age_lena, 
      geom = "histogram",
      binwidth = .1)

qplot(lena_18mo_sym_df$duration, 
      geom = "histogram",
      binwidth = .5)

qplot(lena_18mo_sym_df$awc_prop, 
      geom = "histogram",
      binwidth = .05)

qplot(lena_18mo_sym_df$ctc_prop, 
      geom = "histogram",
      binwidth = .05)

qplot(lena_18mo_sym_df$cvc_prop, 
      geom = "histogram",
      binwidth = .05)

qplot(lena_18mo_sym_df$ctc_daily_rate, 
      geom = "histogram",
      binwidth = 10)

qplot(lena_18mo_sym_df$awc_daily_rate, 
      geom = "histogram",
      binwidth = 50)

qplot(lena_18mo_sym_df$cvc_daily_rate, 
      geom = "histogram",
      binwidth = 10)

count(lena_18mo_sym_df, day_type)

```

# Look at means, std, min, max for variables of interest
```{r}
#in complete dataset
summaries_continuous <-
lena_18mo_sym_df %>% 
  summarise_at(
    vars(
      itsea_competence, 
      itsea_symptoms, 
      crisys_total_6mo_win, 
      crisys_total_6mo,
      awc_prop, 
      ctc_prop, 
      cvc_prop,
      ctc_daily_rate,
      awc_daily_rate,
      cvc_daily_rate
      ), 
    funs(
        mean, 
        sd, 
        min, 
        max), 
    na.rm = TRUE)

#in subset with complete data
summaries_continuous <-
lena_18mo_sym_df_complete %>% 
  summarise_at(
    vars(
      itsea_competence, 
      itsea_symptoms, 
      crisys_total_6mo_win, 
      crisys_total_6mo,
      awc_prop, 
      ctc_prop, 
      cvc_prop,
      ctc_daily_rate,
      awc_daily_rate,
      cvc_daily_rate
      ), 
    list(
        ~mean, 
        ~sd, 
        ~min, 
        ~max,
        ~median), 
    na.rm = TRUE)

```

# Correlation tables
```{r}
#lucy's correlation function
correlations_continuous <-
  lena_18mo_sym_df %>% 
  select(
    itsea_symptoms,
    itsea_competence,
    baby_age_18month,
    baby_age_lena,
    mom_age_6mo_visit,
    income_needs,
    crisys_neg_total_6mo_win,
    duration,
    awc_daily_rate,
    ctc_daily_rate,
    cvc_daily_rate,
    awc_prop,
    ctc_prop,
    cvc_prop,
    #num_stress_total,
    #num_stress_preg,
    #max_sev,
    #sum_sev,
    SUR,
    NEG
  )

corr_table <- correlate(correlations_continuous)

corr_mat <- rcorr(as.matrix(correlations_continuous), type = c("pearson"))
corr_mat

# flattenCorrMatrix <- function(cormat, pmat) {
#   ut <- upper.tri(cormat)
#   data.frame(
#     row = rownames(cormat)[row(cormat)[ut]],
#     column = rownames(cormat)[col(cormat)[ut]],
#     cor  =(cormat)[ut],
#     p = pmat[ut]
#     )
# }
# 
# flattenCorrMatrix(corr_mat$r, corr_mat$p)

#Insignificant correlations are left blank
corrplot(
  corr_mat$r,
  type = "upper", order = "hclust",
  p.mat = corr_mat$P, sig.level = 0.05, insig = "blank"
)


```
# Center continuous variables & median split stress variables

```{r}
# create variables that are median splits for stress variables
lena_18mo_sym_df <-
  lena_18mo_sym_df %>% 
  mutate(
    crisys_median_split = if_else(
      crisys_neg_total_6mo_win > median(crisys_neg_total_6mo_win, na.rm = TRUE),
      "stressed", "not stressed"
    ),
    num_stress_total_median_split = if_else(
      num_stress_total > median(num_stress_total, na.rm = TRUE),
      "high lifetime stress", "low lifetime stress"
    ),
    sumsev_total_median_split = if_else(
      sum_sev > median(sum_sev, na.rm = TRUE),
      "intense stress", "not intense stress"
    )
  )

#create factor with crisys_median_split variables
lena_18mo_sym_df <-
  lena_18mo_sym_df %>% 
  mutate(
    crisys_median_split_num = if_else(
      crisys_median_split == "stressed", 
      1, 0
    ),
    crisys_median_factor = as.factor(crisys_median_split_num)
  )

contrasts(lena_18mo_sym_df$crisys_median_factor) = c(-1,1)

#center continuous variables
lena_18mo_sym_df <-
  lena_18mo_sym_df %>% 
  mutate(
    awc_prop_centered = scale(awc_prop, scale = FALSE),
    itsea_symptoms_centered = scale(itsea_symptoms, scale = FALSE),
    ctc_prop_centered = scale(ctc_prop, scale = FALSE),
    cvc_prop_centered = scale(cvc_prop, scale = FALSE),
    itsea_competence_centered = scale(itsea_competence, scale = FALSE)
  )
```
#test associations among categorical variables 
```{r}
#categorical variables = crisys_median_split, education, male, day_type

##Is day_type related to daily rates, props, symptoms, competence?

lena_18mo_sym_df %>% 
  count(day_type)
#49 weekdays, 21 weekends

t.test(lena_18mo_sym_df$awc_daily_rate ~ lena_18mo_sym_df$day_type)
#ns
t.test(lena_18mo_sym_df$ctc_daily_rate ~ lena_18mo_sym_df$day_type)
#ns
t.test(lena_18mo_sym_df$cvc_daily_rate ~ lena_18mo_sym_df$day_type)
#higher cvc daily rate in weekday vs. weekend group
t.test(lena_18mo_sym_df$awc_prop ~ lena_18mo_sym_df$day_type)
#ns
t.test(lena_18mo_sym_df$ctc_prop ~ lena_18mo_sym_df$day_type)
#ns
t.test(lena_18mo_sym_df$cvc_prop ~ lena_18mo_sym_df$day_type)
#ns
t.test(lena_18mo_sym_df$itsea_symptoms ~ lena_18mo_sym_df$day_type)
#ns
t.test(lena_18mo_sym_df$itsea_competence ~ lena_18mo_sym_df$day_type)
#ns

```
# Main effects
```{r}
#lena variables on itsea symptoms
lena_18mo_sym_awcprop <- 
  lm(
    itsea_symptoms_centered ~ 
      awc_prop_centered,
    data = lena_18mo_sym_df 
                 )

summary(lena_18mo_sym_awcprop)
#ns

lena_18mo_sym_ctcprop <- 
  lm(
    itsea_symptoms ~ 
      ctc_prop_centered,
    data = lena_18mo_sym_df)

summary(lena_18mo_sym_ctcprop)
#ns 

lena_18mo_sym_cvcprop <- 
  lm(
    itsea_symptoms ~ 
      cvc_prop_centered,
    data = lena_18mo_sym_df)
summary(lena_18mo_sym_cvcprop)
#ns

lena_18mo_sym_awcrate <- 
  lm(
    itsea_symptoms ~ 
      awc_daily_rate,
    data = lena_18mo_sym_df)
summary(lena_18mo_sym_awcrate)
#ns

lena_18mo_sym_ctcrate <- 
  lm(
    itsea_symptoms ~ 
      ctc_daily_rate,
    data = lena_18mo_sym_df)
summary(lena_18mo_sym_ctcrate)
#sig

lena_18mo_sym_cvcrate <- 
  lm(
    itsea_symptoms ~ 
      cvc_daily_rate,
    data = lena_18mo_sym_df)
summary(lena_18mo_sym_cvcrate)

#crisys on itsea symptoms

lena_18mo_sym_crisys <- 
  lm(
    itsea_symptoms ~ 
     crisys_neg_total_6mo_win,
    data = lena_18mo_sym_df)
summary(lena_18mo_sym_crisys)
#sig

#visualize significant effects
lena_18mo_sym_df %>%
  ggplot(aes(crisys_neg_total_6mo_win, itsea_symptoms)) +
  geom_point() +
  geom_smooth(method ="lm")

lena_18mo_sym_df %>%
  ggplot(aes(ctc_daily_rate, itsea_symptoms)) +
  geom_point() +
  geom_smooth(method ="lm")
```

#Run main analyses: lena variables * crisys

```{r}
lena_18mo_sym_int_awcprop_neg <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * awc_prop_centered,
    data = lena_18mo_sym_df 
  )

summary(lena_18mo_sym_int_awcprop_neg)

lena_18mo_sym_int_awcrate_neg <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * awc_daily_rate,
    data = lena_18mo_sym_df 
  )

summary(lena_18mo_sym_int_awcrate_neg)


lena_18mo_sym_int_ctcprop_neg <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_prop_centered,
    data = lena_18mo_sym_df 
  )

summary(lena_18mo_sym_int_ctcprop_neg)


lena_18mo_sym_int_ctcrate_neg <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate,
    data = lena_18mo_sym_df 
  )

summary(lena_18mo_sym_int_ctcrate_neg)

```
#calculate R2 change from just interaction term in crisys x ctcrate --> itsea symptoms
```{r}
#model with all predictors separate
lena_18mo_sym_int_ctcrate_neg <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win + ctc_daily_rate,
    data = lena_18mo_sym_df 
  )

r2_1 <- summary(lena_18mo_sym_int_ctcrate_neg)$r.squared

#model with interaction (X = interaction)
lena_18mo_sym_int_ctcrate_negX <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate,
    data = lena_18mo_sym_df 
  )
summary(lena_18mo_sym_int_ctcrate_negX)

r2_2 <- summary(lena_18mo_sym_int_ctcrate_negX)$r.squared

#change from just predictors to predictors & interaction term
r2_change <- r2_2 - r2_1
```
#models just within weekday recordings
```{r}
#create df with just weekdays
lena_18mo_sym_df_weekdays <-
  lena_18mo_sym_df %>% 
  filter(
    day_type == "weekday"
  )

lena_18mo_sym_int_awcprop_neg_weekdays <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * awc_prop_centered,
    data = lena_18mo_sym_df_weekdays 
  )

summary(lena_18mo_sym_int_awcprop_neg_weekdays)

lena_18mo_sym_int_awcrate_neg_weekdays <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * awc_daily_rate,
    data = lena_18mo_sym_df_weekdays 
  )

summary(lena_18mo_sym_int_awcrate_neg_weekdays)


lena_18mo_sym_int_ctcprop_neg_weekdays <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_prop_centered,
    data = lena_18mo_sym_df_weekdays 
  )

summary(lena_18mo_sym_int_ctcprop_neg_weekdays)


lena_18mo_sym_int_ctcrate_neg_weekdays <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate,
    data = lena_18mo_sym_df_weekdays
  )

summary(lena_18mo_sym_int_ctcrate_neg_weekdays)
```

#simple slopes:  crisys * ctcrate interaction 
```{r}
#create new variables: high and low ctc_daily_rate
lena_18mo_sym_df <-
  lena_18mo_sym_df %>% 
  mutate(
    ctc_daily_rate_hi = scale(ctc_daily_rate, scale = F) - sd(ctc_daily_rate),
    ctc_daily_rate_low = scale(ctc_daily_rate, scale = F) + sd(ctc_daily_rate)
    # ctc_prop_hi = scale(ctc_prop, scale = F) - sd(awc_prop),
    # ctc_prop_low = scale(ctc_prop, scale = F) + sd(awc_prop)
  )


#run model with ctc_daily_rate_hi
lena_18mo_sym_ctcrate_hi <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate_hi,
    data = lena_18mo_sym_df 
  )

summary(lena_18mo_sym_ctcrate_hi)
#

#run model with ctc_daily_rate_low
lena_18mo_sym_ctcrate_low <- 
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate_low,
    data = lena_18mo_sym_df 
  )

summary(lena_18mo_sym_ctcrate_low)

```
#visualize interactions using crisys_median_split
```{r}

count(lena_18mo_sym_df, crisys_median_split)

#lena prop variables
lena_18mo_sym_df %>%
  filter(
    !is.na(crisys_median_split)
  ) %>% 
  ggplot(aes(awc_prop_centered, itsea_symptoms, color = crisys_median_split)) +
  geom_point() +
  geom_smooth(method ="lm")

lena_18mo_sym_df %>%
  filter(
    !is.na(crisys_median_split)
  ) %>% 
  ggplot(aes(ctc_prop_centered, itsea_symptoms, color = crisys_median_split)) +
  geom_point() +
  geom_smooth(method ="lm")

#lena daily rate variables
lena_18mo_sym_df %>%
  filter(
    !is.na(crisys_median_split)
  ) %>% 
  ggplot(aes(awc_daily_rate, itsea_symptoms, color = crisys_median_split)) +
  geom_point() +
  geom_smooth(method ="lm")

lena_18mo_sym_df %>%
  filter(
    !is.na(crisys_median_split)
  ) %>% 
  ggplot(aes(ctc_daily_rate, itsea_symptoms, color = crisys_median_split)) +
  geom_point() +
  geom_smooth(method ="lm")

```

##visualizations of ctcrate x crisys --> itsea symptoms 
#1. Simple graph with no error ribbon
#2. Slopes & +1SD, -1SD around mean against ctc rates residualized for baby age at lena
```{r}
#1. simple interaction with non-residualized data
lena_18mo_sym_df %>% 
  ggplot(
    aes(
      crisys_neg_total_6mo_win, 
      itsea_symptoms,
      color = factor(ctc_daily_rate > median(ctc_daily_rate, na.rm = TRUE))
    )
  ) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, size = 2) +
  theme_minimal() +
  labs(
    color = "above ctc_rate median"
  )

#2. make plot with lines = 1SD above, 1SD below mean; points = residuals

#make residuals for symptoms controlling for baby_age_lena
symptoms_residuals_lena_age <-
  lm(
    itsea_symptoms_centered ~ 
      baby_age_lena,
    data = lena_18mo_sym_df 
  )

add_residuals(lena_18mo_sym_df, symptoms_residuals_lena_age, var = "resid_lena_age")

#setting intercept and slope
modsummary_lo <- summary(lena_18mo_sym_ctcrate_low)
intercept_lo <- modsummary_lo$coefficients[1]
slope_lo <- modsummary_lo$coefficients[3]

modsummary_hi <- summary(lena_18mo_sym_ctcrate_hi)
intercept_hi <- modsummary_hi$coefficients[1]
slope_hi <- modsummary_hi$coefficients[3]

simple_slopes_ctcdailyrate <-
  tribble(
    ~effect, ~intercept, ~slope,
    "lower", intercept_lo, slope_lo,
    "higher", intercept_hi, slope_hi
  )

#graph
lena_18mo_sym_df %>% 
  ggplot(aes(crisys_neg_total_6mo, itsea_symptoms_centered)) +
  geom_jitter(alpha = 1/2, size = 3) +
  geom_abline(
    data = simple_slopes_ctcdailyrate,
    aes(
      intercept = intercept,
      slope = slope,
      color = factor(effect)
    ),
    size = 3
  ) +
  scale_color_manual(
    labels = c("+ 1SD", "-1SD"),
    values = c("purple", "orange")
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  ) +
  labs(
    title = "Slope of 18 mo sx by crisys score at\n lower and higher levels of daily ctc",
    color = "CTC daily rate",
    x = "CRISYS score",
    y = "18 month symptoms"
  )

```

#Regression diagnostics - focusing on most significant interaction (crisys * ctcrate)
```{r}
#assessing outliers
outlierTest(lena_18mo_sym_int_ctcrate_neg) # Bonferonni p-value for most extreme obs
qqPlot(lena_18mo_sym_int_ctcrate_neg, main="QQ Plot") #qq plot for studentized resid 
leveragePlots(lena_18mo_sym_int_ctcrate_neg) # leverage plots

# Normality of Residuals
# qq plot for studentized resid
qqPlot(lena_18mo_sym_int_ctcrate_neg, main="QQ Plot")
# distribution of studentized residuals

# Evaluate Collinearity
vif(lena_18mo_sym_int_ctcrate_neg) # variance inflation factors 
sqrt(vif(lena_18mo_sym_int_ctcrate_neg)) > 2 

# Global test of model assumptions
library(gvlma)
gvmodel <- gvlma(lena_18mo_sym_int_ctcrate_neg) 
summary(lena_18mo_sym_int_ctcrate_neg)
```

#olsrr diagnostics, standardized betas: ctcrate * crisys
```{r}
##standardized betas

library(olsrr)
ols_regress_model_middle <-
ols_regress(itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate,
    data = lena_18mo_sym_df)
print(ols_regress_model_middle)
#ols_plot_diagnostics(lena_18mo_sym_int_awc) - not really sure why this isn't working
 #standardized betas: interaction term: -.341 (-.317, -.044); awc_prop_centered: -.150 (-.287, 1.655); crisys_total_6mo_win: .237 (-.001, .027)

ols_regress_model_high <-
ols_regress(itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate_hi,
    data = lena_18mo_sym_df)
print(ols_regress_model_high)
 
ols_regress_model_low <-
ols_regress(itsea_symptoms_centered ~ 
      baby_age_lena + crisys_neg_total_6mo_win * ctc_daily_rate_low,
    data = lena_18mo_sym_df)
print(ols_regress_model_low)


##olsrr diagnostics
plot(lena_18mo_sym_int_awcrate_neg)

```
#examine LENA follow-up data
```{r}
#Within all ps with followup data and awc_prop
lena_followup_and_data <-
  read_csv(lena_follow_and_data_file)

#see if percent_mother associated with ctc_daily_rate
ctc_daily_rate <- 
  lm(
    ctc_daily_rate ~ percent_mother,
    data = lena_followup_and_data 
    )
summary(ctc_daily_rate)
#association not significant (p = .69)

#visualize regression
lena_followup_and_data %>% 
ggplot(aes(percent_mother, ctc_daily_rate)) +
  geom_point() +
  geom_smooth(method ="lm")

#report stats on percent_mother
lena_followup_and_data %>% 
  summarize_at(
    vars(
    percent_mother
    ), 
    list(~mean, ~sd, ~min, ~max),
    na.rm = TRUE)

#see if percent_mother associated with weekday vs weekend
t.test(lena_followup_and_data$percent_mother ~ lena_followup_and_data$day_type)
  
```

#save dataframe used for analyses 
```{r}
#this is a check to make sure nothing looks weird
write_csv(lena_18mo_sym_df, "~/Desktop/lena_18mo_symptoms/data_for_analyses_20190510.csv")
```
```

