---
title: "Analyses for _Dimensions of the language environment in infancy and symptoms of psychopathology in toddlerhood_"
author: "Lucy S. King"
date: "2020"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

# Environment
```{r}
# Libraries
library(tidyverse)
library(lubridate)
library(corrr)
library(corrplot)
library(ggpubr)
library(modelr)
library(effectsize)
library(glmpath)
library(glmnet)
library(MASS)
library(ggcorrplot)
library(broom)
library(ggfortify)
library(relaimpo)
library(olsrr)
library(parameters)
library(BayesFactor)
library(codebook)
library(labelled)


# Files 
home <- "~/Box/Mooddata_Coordinating/BABIES/Data/final_scored_data/"

eligible_id_file <- "~/Box/lucy_king_files/BABIES/lena_symptoms/inclusion_tracker_sub2.xlsx"
lena_edited_file <- "~/Box/lucy_king_files/BABIES/lena_symptoms/included_recording_dates.csv"
fp_intervention_id <- paste0(home, "lab_caregiving_behavior/free_play_intervention_assignment.csv")
itsea_file <- paste0(home, "survey_18month/itsea_scored_20201001.csv")
demographics_file <- paste0(home, "demographics/demo_6mo_cleaned_final.csv")
lena_file <- paste0(home, "LENA/lena_final_wf_day1_20200208_upto9months.csv")
ibq_file <- paste0(home, "IBQ/ibq_scored_final.csv")
sfp_care_file <- paste0(home, "lab_caregiving_behavior/PCIRS_sfp_wf_complete.csv")
fp_care_file <- paste0(home, "lab_caregiving_behavior/free_play_wf_8min_20200507.csv")
cesd_file <- paste0(home, "CESD/cesd_wf_20201001.csv")

# Functions
source("identify_outliers_histogram.R")

theme_lena <-
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  )

# Display options
options(scipen = 999)
```


# Read in and join data
```{r}
d0 <-
  readxl::read_xlsx(eligible_id_file, sheet = "data_filtered_T3") %>% 
  dplyr::select(
    ID,
    eligible,
    included
  ) %>% 
  filter(included == 1) %>% 
  left_join(
    read_csv(lena_edited_file) %>% 
      dplyr::select(
        ID, LENA_edited
      ),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(demographics_file) %>% 
      dplyr::select(
        ID,
        baby_dob,
        education,
        education_txt,
        annual_income_txt,
        income_numeric = annual_income_num,
        ppl_in_home_total = ppl_in_home_tot,
        ppl_in_home_allchild,
        income_needs,
        male,
        mom_age,
        mom_latinx,
        mom_race,
        age_behav,
        secondlang,
        secondlang_type
      ),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(itsea_file) %>% 
      dplyr::select(
        ID,
        survey_date,
        itsea_symptoms,
        itsea_activity,
        itsea_agress,
        itsea_gad,
        itsea_negemo,
        itsea_depp,
        itsea_intern,
        itsea_extern,
        # itsea symptoms items
        itsea_a2, 
        itsea_a4, 
        itsea_b8, 
        itsea_b10, 
        itsea_b27, 
        itsea_b32,
        itsea_a28, 
        itsea_a30, 
        itsea_a33, 
        itsea_b3, 
        itsea_b9, 
        itsea_b16, 
        itsea_b30, 
        itsea_b34, 
        itsea_b44, 
        itsea_b88, 
        itsea_c2, 
        itsea_e4,
        itsea_a9, 
        itsea_a39, 
        itsea_a40, 
        itsea_b43, 
        itsea_b76, 
        itsea_b81, 
        itsea_b84, 
        itsea_b91, 
        itsea_b92,
        itsea_a3, 
        itsea_a12, 
        itsea_a35, 
        itsea_b37, 
        itsea_b70, 
        itsea_b86, 
        itsea_e5, 
        itsea_e6, 
        itsea_e11,
        itsea_a7, 
        itsea_a21, 
        itsea_a23, 
        itsea_b31, 
        itsea_b45, 
        itsea_b50, 
        itsea_b53, 
        itsea_b59, 
        itsea_b65, 
        itsea_b66, 
        itsea_b74, 
        itsea_b80, 
        itsea_b85,
        itsea_concern_ied
      ), 
    by = "ID"
  ) %>% 
  mutate(
    survey_date =  parse_date_time(survey_date, orders = c("mdy HM")),
    age_18mo = (baby_dob %--% survey_date) / months(1)
  ) %>% 
  left_join(
    read_csv(fp_care_file) %>% 
      dplyr::select(ID, negmood_FP, sens_FP),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(sfp_care_file) %>% 
      dplyr::select(ID, sens_R_M, sens_M),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(ibq_file) %>% 
      dplyr::select(
        ID,
        #IBQ NEG items
        ibq_64, ibq_74, ibq_75, ibq_32, ibq_79, ibq_80,
        ibq_2, ibq_3_r, ibq_4, ibq_21, ibq_52, ibq_53, ibq_62,
        ibq_22, ibq_76, ibq_77, ibq_78, ibq_87, ibq_89,
        ibq_36, ibq_37, ibq_38, ibq_63, ibq_71, ibq_72,
        NEG
      ),
    by = "ID"
  ) %>% 
  left_join(read_csv(lena_file), by = "ID") %>% 
  left_join(read_csv(cesd_file), by = "ID") %>% 
  left_join(read_csv(fp_intervention_id), by = "ID") %>% 
  rename(
    age_lena_d1 = age,
    sens_r_sfp = sens_R_M,
    sens_sfp = sens_M,
    sens_fp = sens_FP,
    negmood_fp = negmood_FP
  ) 
```

# Calculate income-to-needs based on 200% of federal poverty line limits in 2017
```{r}
d0 <-
  d0 %>% 
  group_by(ID) %>% 
  mutate(
    # inr based on the federal poverty thresholds for number of people in household and number that are children
    fpl_threshold = case_when(
      ppl_in_home_total == 2 ~ 16493,
      ppl_in_home_total == 3 ~ 19515,
      ppl_in_home_total == 4 ~ 25094,
      ppl_in_home_total == 5 ~ 29714,
      ppl_in_home_total == 6 ~ 33618,
      ppl_in_home_total == 7 ~ 38173,
      ppl_in_home_total == 8 ~ 42684,
      ppl_in_home_total >= 9 ~ 50681
    ),
    inr_fpl = income_numeric / fpl_threshold,
    fpl_low_income = inr_fpl < 1
  ) %>% 
  ungroup()
```

# Retention / Missing data

Dyads who completed the 6-month assessment (either lab or home observation), were eligible for the 18-month assessment (i.e., were age 18 months by the time of analysis), and completed the 18-month assessment are included in the current analyses. 

```{r}
#missing SFP
d0 %>% 
  count(is.na(sens_sfp))

#missing FP
d0 %>% 
  count(is.na(sens_fp))

#missing IBQ
d0 %>% 
  count(is.na(NEG))

d0 %>% 
  dplyr::select(
    awc_prop, 
    ctc_prop,
    awc_hour_max,
    ctc_hour_max,
    sens_r_sfp,
    sens_fp,
    NEG,
    cesd_t4,
    itsea_symptoms
  ) %>% 
  pair_n()
```


# Descriptive statistics 

## IBQ, ITSEA reliability

```{r}
d0 %>% 
  dplyr::select(ibq_64:ibq_72) %>% 
  psych::alpha(check.keys = TRUE)
```

```{r}
d0 %>% 
  dplyr::select(itsea_a2:itsea_b85, -itsea_c2) %>% 
  psych::alpha()
```

```{r}
d0 %>% 
  summarize_at(
    vars(
      age_18mo,
      age_lena_d1,
      NEG,
      duration_total,
      duration_analyzed,
      ct_first_time,
      ct_last_time,
      cvc_hour_max,
      ctc_hour_max,
      awc_hour_max,
      ctc_prop,
      awc_prop,
      sens_r_sfp,
      sens_fp,
      negmood_fp,
      itsea_symptoms,
      mom_age,
      income_needs,
      inr_fpl,
      percent_mother,
      cesd_t3,
      cesd_t4
    ), 
    funs(mean, sd, min, max), na.rm = TRUE
  )
```

```{r}
d0 %>% 
  count(male) %>% 
  mutate(
    per = n / sum(n)
  )

d0 %>% 
  count(itsea_concern_ied) %>% 
  mutate(
    per = n / sum(n)
  )

d0 %>% 
  count(income_needs < 1) %>% 
  mutate(
    per = n / sum(n)
  )

d0 %>% 
  count(inr_fpl < 1) %>% 
  mutate(
    per = n / sum(n)
  )

d0 %>% 
  count(income_numeric <= 2 * fpl_threshold) %>% 
  mutate(
    per = n / sum(n)
  )

d0 %>% 
  count(mom_latinx) %>% 
  mutate(
    per = n / sum(n)
  )

d0 %>% 
  count(mom_race) %>% 
  mutate(
    per = n / sum(n)
  ) %>% 
  arrange(desc(n))

d0 %>% 
  count(education_txt) %>% 
  mutate(
    per = n / sum(n)
  ) %>% 
  arrange(education_txt)

d0 %>% 
  count(day_type) %>% 
  mutate(per = n / sum(n))

d0 %>% 
  count(cesd_t3 >= 16) %>% 
  mutate(per = n / sum(n))

d0 %>% 
  count(cesd_t4 >= 16) %>% 
  mutate(per = n / sum(n))

```

```{r}
d0 %>% 
  count(lena_mornstart, lena_recordfull)

d0 %>% 
  filter(lena_mornstart == 0) %>% 
  dplyr::select(
    ID, 
    lena_recordno, 
    ct_first_time, 
    ct_last_time, 
    duration_total, 
    duration_analyzed, 
    lena_notes
  )
```

## Distributions 

```{r}
d0 %>% 
  identify_outliers_hist(x = cvc_hour_max)

d0 %>% 
  identify_outliers_hist(x = ctc_hour_max)

d0 %>% 
  identify_outliers_hist(x = ctc_prop)

d0 %>% 
  identify_outliers_hist(x = awc_hour_max)

d0 %>% 
  identify_outliers_hist(x = awc_prop)

d0 %>% 
  identify_outliers_hist(x = sens_r_sfp)

d0 %>% 
  identify_outliers_hist(x = sens_fp)

d0 %>% 
  identify_outliers_hist(x = itsea_symptoms)

d0 %>% 
  identify_outliers_hist(x = cesd_t3)

d0 %>% 
  identify_outliers_hist(x = cesd_t4)


d0 %>% 
  ggplot(aes(percent_mother)) +
  geom_histogram() +
  theme_lena +
  labs(
    x = "% of the LENA recording day infant with mother"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/percent_mom_hist.png",
  width = 7,
  height = 5
)

d0 %>% 
  ggplot(aes(income_needs)) +
  geom_histogram(binwidth = .5) +
  geom_vline(
    xintercept = 1,
    color = "red"
  ) +
  scale_x_continuous(breaks = seq.int(0, 6, .5)) +
  theme_lena +
  labs(
    x = "Income-to-needs ratio based on\n county-specifc HUD low-income limits"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/inr_HUD_hist.png",
  width = 7,
  height = 5
)

d0 %>% 
  ggplot(aes(inr_fpl)) +
  geom_histogram(binwidth = 1) +
  geom_vline(
    xintercept = 1,
    color = "red"
  ) +
  scale_x_continuous(breaks = seq.int(0, 6, 1)) +
  theme_lena +
  labs(
    x = "Income-to-needs ratio based on\n federal poverty line limits"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/inr_FPL_hist.png",
  width = 7,
  height = 5
)
```


## Correlations

``` {r}
corr_lena <- 
  d0 %>%
  dplyr::select(
    `AW consistency` = awc_prop,
    `CT consistency` = ctc_prop,
    `AW quantity` = awc_hour_max,
    `CT quantity` = ctc_hour_max,
    `Infant vocalizations` = cvc_hour_max,
    `Sensitivity to distress` = sens_r_sfp,
    `Sensitivity during play` = sens_fp,
  ) %>% 
  correlate(use = "pairwise.complete.obs", method = "pearson") %>% 
  shave(upper = TRUE) %>% 
  fashion()  
```

```{r}
corr_lena_plot <-
  corr_lena %>% 
  gather(variable, value, -rowname) %>% 
  mutate(
    variable = str_replace_all(variable, "\\.", " "),
    rowname = as.character(rowname),
    value_chr = as.character(value),
    value_chr = case_when(
      rowname == "CT consistency" & variable == "AW quantity" ~ ".23",
      rowname == "AW quantity" & variable == "CT consistency" ~ NA_character_,
      rowname == "Sensitivity to distress" & variable == "Sensitivity during play" ~ ".10",
      rowname == "Sensitivity during play" & variable == "Sensitivity to distress" ~ NA_character_,
      TRUE ~ value_chr
    ),
    value_num = as.numeric(value_chr)
  ) 
```

```{r}
corr_lena_plot %>% 
  ggplot(aes(x = rowname, y = variable)) +
  geom_tile(aes(fill = abs(value_num))) +
  geom_text(
    aes(label = value_chr), 
    size = 4
  ) +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    na.value = "white",
    midpoint = 0,
    limit = c(-1, 1),
    space = "Lab",
    name = "Pearson correlation\ncoefficient"
  ) +
  theme_void() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1,
      size = 12
    ),
    axis.text.y = element_text(
      size = 12,
      hjust = 1.1
    )
  ) 

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/lena_corr_plot.eps", 
  dpi = 1000,
  height = 5, 
  width = 7
)
```

```{r}
cor.test(d0$cesd_t3, d0$awc_hour_max)
cor.test(d0$cesd_t3, d0$awc_prop)
cor.test(d0$cesd_t3, d0$ctc_hour_max)
cor.test(d0$cesd_t3, d0$ctc_prop)
cor.test(d0$cesd_t3, d0$cvc_hour_max)
cor.test(d0$cesd_t3, d0$sens_r_sfp)
cor.test(d0$cesd_t3, d0$sens_fp)

cor.test(d0$NEG, d0$awc_hour_max)
cor.test(d0$NEG, d0$awc_prop)
cor.test(d0$NEG, d0$ctc_hour_max)
cor.test(d0$NEG, d0$ctc_prop)
cor.test(d0$NEG, d0$cvc_hour_max)
cor.test(d0$NEG, d0$sens_r_sfp)
cor.test(d0$NEG, d0$sens_fp)
```

```{r}
cor.test(d0$awc_hour_max, d0$sens_fp)
cor.test(d0$awc_hour_max, d0$sens_r_sfp)

cor.test(d0$ctc_hour_max, d0$sens_fp)
cor.test(d0$ctc_hour_max, d0$sens_r_sfp)

cor.test(d0$awc_prop, d0$sens_fp)
cor.test(d0$awc_prop, d0$sens_r_sfp)

cor.test(d0$ctc_prop, d0$sens_fp)
cor.test(d0$ctc_prop, d0$sens_r_sfp)

cor.test(d0$cvc_hour_max, d0$sens_fp)
cor.test(d0$cvc_hour_max, d0$sens_r_sfp)

cor.test(d0$sens_fp, d0$sens_r_sfp)
```

## t-tests by bilingual status
```{r}
t.test(d0$awc_hour_max ~ d0$secondlang)

t.test(d0$ctc_hour_max ~ d0$secondlang)

t.test(d0$ctc_prop ~ d0$secondlang)

t.test(d0$awc_prop ~ d0$secondlang)

t.test(d0$itsea_symptoms ~ d0$secondlang)
```

## t-tests by day type
```{r}
t.test(d0$awc_hour_max ~ d0$day_type)

t.test(d0$ctc_hour_max ~ d0$day_type)

t.test(d0$ctc_prop ~ d0$day_type)

t.test(d0$awc_prop ~ d0$day_type)
```


## Histograms
```{r}
d0 %>% 
  dplyr::select(
    ctc_hour_max,
    awc_hour_max,
    ctc_prop,
    awc_prop,
    sens_fp,
    sens_r_sfp
  ) %>% 
  gather(measure, value, ctc_hour_max:sens_r_sfp) %>% 
  mutate(
    measure = factor(
      measure,
      levels = c(
        "awc_prop",
        "ctc_prop",
        "awc_hour_max",
        "ctc_hour_max",
        "sens_r_sfp",
        "sens_fp"
      ),
      labels = c(
        "AW consistency", 
        "CT consistency", 
        "AW quantity", 
        "CT quantity",
        "Sensitivity to distress",
        "Sensivity during play"
        )
    )
  ) %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(.~measure, scales = "free", ncol = 2, nrow = 3) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 18),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Value"
  )


ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/care_hist_ggpanels.png",
  width = 7.5,
  height = 8
)
```

```{r}
d0 %>% 
  ggplot(aes(itsea_symptoms)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq.int(0, 3, .25)) +
  theme_lena +
  labs(
    x = "Toddler symptoms of psychopathology"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/itsea_hist.png"
)
```

```{r}
d0 %>% 
  dplyr::select(
    Activity = itsea_activity,
    Aggression = itsea_agress,
    Depression = itsea_depp,
    Anxiety = itsea_gad,
    `Negative emotion` = itsea_negemo
  ) %>% 
  gather(itsea_subscale, score, Activity:`Negative emotion`) %>% 
  ggplot(aes(score)) +
  geom_histogram(bins = 12) +
  theme_lena +
  theme(
    strip.text = element_text(size = 16)
  ) +
  facet_wrap(.~itsea_subscale, scales = "free") +
  labs(
    x = "Subscale score"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/itsea_subscale_hist.png",
  width = 10,
  height = 7
)
```

```{r}
d0 %>% 
  ggplot(aes(cesd_t3)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq.int(0, 50, 5)) +
  theme_lena +
  labs(
    x = "Maternal depressive symptoms\n at infant age 6 months"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/cesd6_hist.png"
)
```

```{r}
d0 %>% 
  ggplot(aes(cesd_t4)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq.int(0, 50, 5)) +
  theme_lena +
  labs(
    x = "Maternal depressive symptoms\n at infant age 18 months"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/cesd18_hist.png"
)
```

# Results 

## Associations of LENA measures with ITSEA symptoms 

### Effect code categorical variables

```{r}
d0 <-
  d0 %>% 
  mutate(
    white = as.factor(
      if_else(
        mom_race == "White", 
        1, 0
      )
    ),
    male = as.factor(male),
    mom_latinx = as.factor(mom_latinx)
  )

contrasts(d0$male) <- c(-.5, .5)
contrasts(d0$white) <- c(-.5, .5)
contrasts(d0$mom_latinx) <- c(-.5, .5)
```

### Formal model fitting
```{r}
d0_lm <-
  d0 %>% 
  filter(!is.na(NEG))
```


```{r}
mod_0 <-
  lm(
    scale(itsea_symptoms) ~
      1,
    data = d0_lm
  )
```

```{r}
mod_cesdt3 <-
  lm(
    scale(itsea_symptoms) ~
      scale(cesd_t3),
    data = d0_lm
  )

anova(mod_0, mod_cesdt3)
```

```{r}
mod_cesdt4 <-
  lm(
    scale(itsea_symptoms) ~
      scale(cesd_t4),
    data = d0_lm
  )

anova(mod_0, mod_cesdt4)
```

```{r}
mod_neg <-
  lm(
    scale(itsea_symptoms) ~
      scale(cesd_t4) +
      scale(NEG),
    data = d0_lm
  )

anova(mod_cesdt4, mod_neg)
model_parameters(mod_neg)
calc.relimp(mod_neg)
```

```{r}
mod_Bage <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(age_lena_d1),
    data = d0_lm
  )

anova(mod_neg, mod_Bage)
```

```{r}
mod_Mage <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(mom_age),
    data = d0_lm
  )

anova(mod_neg, mod_Mage)
```

```{r}
mod_race <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      white, 
    data = d0_lm
  )

anova(mod_neg, mod_race)
```

```{r}
mod_ethnicity <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      mom_latinx, 
    data = d0_lm
  )

anova(mod_neg, mod_ethnicity)
```

```{r}
mod_education <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(education),
    data = d0_lm
  )

anova(mod_neg, mod_education)
```

```{r}
mod_income_needs <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(income_needs), 
    data = d0_lm
  )

anova(mod_neg, mod_income_needs)
```
```{r}
mod_inr_fpl <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(inr_fpl), 
    data = d0_lm
  )

anova(mod_neg, mod_inr_fpl)
```

```{r}
mod_sex <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      male, 
    data = d0_lm
  )

anova(mod_neg, mod_sex)
```
```{r}
mod_duration <-
    lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(duration_analyzed),
    data = d0_lm
  )

anova(mod_neg, mod_duration)
```

### Consistency

#### AW
```{r}
mod_awprop <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(awc_prop),
    data = d0_lm
  )

model_parameters(mod_awprop)
calc.relimp(mod_awprop)
```

#### CT
```{r}
mod_ctprop <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(ctc_prop),
    data = d0_lm
  )

model_parameters(mod_ctprop)
calc.relimp(mod_ctprop)
```

### Quantity

#### AW
```{r}
mod_awmax<-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(awc_hour_max),
    data = d0_lm
  )

model_parameters(mod_awmax)
calc.relimp(mod_awmax)
```
#### CT
```{r}
mod_ctmax <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(ctc_hour_max),
    data = d0_lm
  )

model_parameters(mod_ctmax)
calc.relimp(mod_ctmax)
```

### Infant vocalizations
```{r}
mod_cvmax <-
  lm(
    scale(itsea_symptoms) ~
      scale(NEG) +
      scale(cesd_t4) +
      scale(cvc_hour_max),
    data = d0_lm
  )

model_parameters(mod_cvmax)
calc.relimp(mod_cvmax)
```

### Bayes Factors
```{r}
model_d <-
  d0_lm %>% 
  dplyr::select(
    awc_prop,
    ctc_prop,
    awc_hour_max,
    ctc_hour_max,
    cvc_hour_max,
    itsea_symptoms,
    NEG,
    cesd_t4
  ) %>% 
  as.data.frame()

# each measure is compared to a covariate only model
lmBF_covonly = lmBF(itsea_symptoms ~ NEG + cesd_t4, data = model_d)

lmBF_awcprop = lmBF(itsea_symptoms ~ awc_prop + cesd_t4 + NEG, data = model_d)
lmBF_ctcprop = lmBF(itsea_symptoms ~ ctc_prop + cesd_t4 + NEG, data = model_d)

lmBF_awcmax = lmBF(itsea_symptoms ~ awc_hour_max + cesd_t4 + NEG, data = model_d)
lmBF_ctcmax = lmBF(itsea_symptoms ~ ctc_hour_max + cesd_t4 + NEG, data = model_d)

mBF_cvcmax = lmBF(itsea_symptoms ~ cvc_hour_max + cesd_t4 + NEG, data = model_d)

# BF for AW consistency
lmBF_awcprop / lmBF_covonly

# BF for CT consistency
lmBF_ctcprop / lmBF_covonly

# BF for AW quantity
lmBF_awcmax / lmBF_covonly

# BF for CT quantity
lmBF_ctcmax / lmBF_covonly

# BF for infant vocalizations
mBF_cvcmax / lmBF_covonly
```

### Create residualized symptoms variable for covariates in best fitting model
```{r}
best_cov <- lm(
  itsea_symptoms ~
    scale(cesd_t4, scale = FALSE) +
    scale(NEG, scale = FALSE),
  data = d0
)

d0 <-
  d0 %>% 
  add_residuals(best_cov, var = "resid_symptoms")
```



## Plot associations 

### AW consistency
```{r}
d0 %>% 
  ggplot(aes(awc_prop, resid_symptoms)) +
  geom_point(size = 4, alpha = 1/2) +
  geom_smooth(method = "lm", size = 3, color = "black") +
  scale_x_continuous(
    breaks = seq.int(0, 1, .10),
    labels = scales::percent_format(accuracy = 1)
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 17),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Consistency of adult words in daily life",
    y = "Toddler symptoms of psychopathology\n(residuals)"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/aw_prop_symptoms.png",
  width = 7,
  height = 6
)
```
### CT consistency
```{r}
d0 %>% 
  ggplot(aes(ctc_prop, resid_symptoms)) +
  geom_point(size = 4, alpha = 1/2) +
  geom_smooth(method = "lm", size = 3, color = "black") +
  scale_x_continuous(
    breaks = seq.int(0, 1, .10),
    labels = scales::percent_format(accuracy = 1)
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 17),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Consistency of conversational turns in daily life",
    y = "Toddler symptoms of psychopathology\n(residuals)"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/ctc_prop_symptoms.png",
  width = 7,
  height = 6
)
```
### Consistency 
```{r}
d0 %>% 
  dplyr::select(ctc_prop, awc_prop, resid_symptoms) %>% 
  rename(
    `CT consistency` = ctc_prop,
    `AW consistency` = awc_prop
  ) %>% 
  gather(measure, value, -resid_symptoms) %>% 
  ggplot(aes(value, resid_symptoms)) +
  geom_point(size = 4, alpha = 1/2) +
  geom_smooth(method = "lm", size = 3, color = "black") +
  scale_x_continuous(
    breaks = seq.int(0, 1, .15),
    labels = scales::percent_format(accuracy = 1)
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 17),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Consistency of language input in daily life",
    y = "Toddler symptoms of psychopathology\n(residuals)"
  ) +
  facet_grid(.~measure, scales = "free")

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/ctc_awc_prop_symptoms.eps",
  width = 10,
  height = 6,
  dpi = 1000
)
```

## Infant vocalizations
```{r}
d0 %>% 
  ggplot(aes(cvc_hour_max, resid_symptoms)) +
  geom_point(size = 4, alpha = 1/2) +
  geom_smooth(method = "lm", size = 3, color = "black") +
  scale_x_continuous(
    breaks = seq.int(0, 1000, 100)
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 17),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Quantity of infant vocalizations in daily life",
    y = "Toddler symptoms of psychopathology\n(residuals)"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/cv_max_symptoms.png",
  width = 7,
  height = 6
)
```

## Elastic net regression with LENA variables and covariates

### Prepare data
```{r}
z_score <- function(x) {
  diff_mu <- x - mean(x, na.rm = T)
  sd <- sd(x, na.rm = T)
  diff_mu / sd
}
```

```{r}
d0_en <-
  d0 %>% 
  dplyr::select(
    itsea_symptoms,
    ctc_prop,
    awc_prop,
    ctc_hour_max,
    awc_hour_max,
    cvc_hour_max,
    cesd_t4,
    NEG
  ) %>% 
  mutate_at(
    vars(itsea_symptoms:NEG),
    funs(z_score)
  ) %>% 
  na.omit()
```

```{r}
predictors <- 
  d0_en %>% 
  dplyr::select(-itsea_symptoms) %>% 
  as.matrix()
  
symptoms <- z_score(d0_en$itsea_symptoms)
```

### Run model over 100 increasingly relaxed lambda values
```{r}
fit_net <- glmnet(predictors, symptoms, family = "gaussian", standardize = FALSE)
plot(fit_net, label = TRUE)
```

### Run leave-one-out cross validation to identify optimal lambda value
If nfolds is set to the sample size = leave-one-out CV
https://sciphy-stats.com/post/2019-01-25-finalizing-glmnet-models/
Running the CV 100 times to minimize simulation error. 

```{r}
k <- nrow(d0_en)
```

```{r}
# run cross validation
set.seed(123)

fit_cv <- cv.glmnet(
    predictors, 
    symptoms, 
    type.measure = "mse", 
    family = "gaussian", 
    alpha = .5,
    nfolds = k,
    grouped = FALSE,
    standardized = FALSE
  )

lambda_optimal <- fit_cv$lambda.min
```


### Print coefficients for model with optimal lambda 
```{r}
fit_net_final <- glmnet(
  predictors, 
  symptoms, 
  family = "gaussian", 
  alpha = .5, 
  lambda = lambda_optimal
)
coefs <- coef(fit_net_final) 

en_coef_tbl <- 
  tidy(coef(fit_net_final))  %>% 
  dplyr::select(-column) %>% 
  rename(
    Variable = row,
    `Elastic net estimate` = value
  ) %>% 
  mutate(
    Variable = recode(
      Variable,
      "(Intercept)" = "(Intercept)",
      "ctc_prop" = "CT consistency",
      "awc_prop" = "AW consistency",
      "ctc_hour_max" = "CT quantity",
      "cvc_hour_max" = "Infant vocalizations",
      "cesd_t4" = "Maternal depressive symptoms",
      "NEG" = "Infant negative affectivity"
    )
  ) %>% 
  mutate_at(vars(`Elastic net estimate`) , round, 2)
```
### Calculate zero-order effect sizes
```{r}
cor.test(d0_en$ctc_prop, d0_en$itsea_symptoms)
cor.test(d0_en$awc_prop, d0_en$itsea_symptoms)
cor.test(d0_en$ctc_hour_max, d0_en$itsea_symptoms)
cor.test(d0_en$cvc_hour_max, d0_en$itsea_symptoms)
cor.test(d0_en$cesd_t4, d0_en$itsea_symptoms)
cor.test(d0_en$NEG, d0_en$itsea_symptoms)
```

### Calculate EN model performance
```{r}
# save model matrix
X <- model.matrix(itsea_symptoms ~ ., d0_en)

# get predicted values
y_hat <- matrix(X, ncol = 8) %*% as.matrix(coefs)

# calculate R-squared
r_sq <- 1 - (sum((y_hat[, 1] - symptoms)^2) / sum((symptoms - mean(symptoms))^2))
```

## Additional analyses

### Covarying maternal sensitivity 

#### AW 
```{r}
sym_aw_ms <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(awc_prop) + 
      scale(cesd_t4) +
      scale(NEG) +
      scale(sens_fp) +
      scale(sens_r_sfp),
    data = d0 
  )


calc.relimp(sym_aw_ms)
model_parameters(sym_aw_ms)
```
#### CT
```{r}
sym_ct_ms <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(ctc_prop) +
      scale(cesd_t4) +
      scale(NEG) +
      scale(sens_fp) +
      scale(sens_r_sfp),
    data = d0 
  )

calc.relimp(sym_ct_ms)
model_parameters(sym_ct_ms)
```

```{r}
cor.test(d0$sens_fp, d0$itsea_symptoms)
cor.test(d0$sens_r_sfp, d0$itsea_symptoms)
```

### Who is talking?

```{r}
d0_follow <-
  d0 %>% 
  filter(
    lena_caregivers___1 != 0 |
      lena_caregivers___2 != 0 |
      lena_caregivers___3 != 0 |
      lena_caregivers___4 != 0 |
      lena_caregivers___5 != 0 |
      lena_caregivers___6 != 0 |
      lena_caregivers___7 != 0 |
      lena_caregivers___8 != 0 |
      lena_caregivers___9 != 0 |
      lena_caregivers___10 != 0
  ) %>% 
  rename(
    mom_present = lena_caregivers___1,
    dad_present = lena_caregivers___2,
    daycare_present = lena_caregivers___3,
    nanny_present = lena_caregivers___4,
    sibling_present = lena_caregivers___5,
    cousin_present = lena_caregivers___6,
    grandparent_present = lena_caregivers___7,
    friends_present = lena_caregivers___8,
    relative_present = lena_caregivers___9, #any other relative
    other_present = lena_caregivers___10
  )
```

```{r}
d0_follow %>% 
  count(!is.na(percent_mother))

d0_follow %>%
  summarise_at(
    vars(percent_mother),
    funs(mean, median, sd, min, max), na.rm = TRUE
  )

cor.test(d0_follow$percent_mother, d0_follow$ctc_prop, method = "spearman")
cor.test(d0_follow$percent_mother, d0_follow$awc_prop, method = "spearman")
cor.test(d0_follow$percent_mother, d0_follow$ctc_hour_max, method = "spearman")
cor.test(d0_follow$percent_mother, d0_follow$awc_hour_max, method = "spearman")
cor.test(d0_follow$percent_mother, d0_follow$cvc_hour_max, method = "spearman")

cor.test(d0_follow$percent_mother, d0_follow$itsea_symptoms, method = "spearman")
```

```{r}
d0_follow %>%
  count(mom_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(dad_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(mom_present, dad_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(daycare_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(nanny_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(grandparent_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(friends_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(relative_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_follow %>%
  count(
    mom_present, 
    dad_present, 
    daycare_present, 
    nanny_present, 
    grandparent_present
  ) %>% 
  mutate(
    per = n / sum(n)
  ) %>% 
  arrange(n)

```

```{r}
d0_follow %>% 
  count(percent_mother) %>% 
  arrange(desc(n))
```


```{r}
d0_follow <-
  d0_follow %>% 
  mutate(
    n_caregivers = 
      mom_present +
      dad_present +
      daycare_present +
      grandparent_present +
      nanny_present +
      friends_present +
      relative_present
  )

cor.test(d0_follow$itsea_symptoms, d0_follow$n_caregivers, method = "spearman")
```

### LENA recording correlations: duration, etc. 
```{r}
d0 %>% 
  dplyr::select(
    ctc_prop,
    awc_prop,
    ctc_hour_max,
    awc_hour_max,
    duration_analyzed,
    duration_total,
    ct_first_time,
    ct_last_time
  ) %>% 
  correlate(method = "spearman") %>% 
  fashion()

cor.test(d0$ctc_prop, d0$duration_analyzed, method = "pearson")
cor.test(d0$awc_prop, d0$duration_analyzed, method = "pearson")
cor.test(d0$ctc_hour_max, d0$duration_analyzed, method = "pearson")
cor.test(d0$awc_hour_max, d0$duration_analyzed, method = "pearson")

cor.test(d0$ctc_prop, d0$ct_first_time, method = "pearson")
cor.test(d0$awc_prop, d0$ct_first_time, method = "pearson")
cor.test(d0$ctc_hour_max, d0$ct_first_time, method = "pearson")
cor.test(d0$awc_hour_max, d0$ct_first_time, method = "pearson")

cor.test(d0$ctc_prop, d0$ct_last_time, method = "pearson")
cor.test(d0$awc_prop, d0$ct_last_time, method = "pearson")
cor.test(d0$ctc_hour_max, d0$ct_last_time, method = "pearson")
cor.test(d0$awc_hour_max, d0$ct_last_time, method = "pearson")

cor.test(d0$ct_last_time, d0$ct_first_time, method = "pearson")
```



```{r}
d0 %>% 
  ggplot(aes(duration_analyzed, awc_prop)) +
  geom_point() +
  geom_smooth(method = "lm")

d0 %>% 
  ggplot(aes(ct_first_time, awc_hour_max)) +
  geom_point() +
  geom_smooth(method = "lm")

d0 %>% 
  ggplot(aes(ct_first_time, ctc_hour_max)) +
  geom_point() +
  geom_smooth(method = "lm")

d0 %>% 
  ggplot(aes(ct_last_time, awc_prop)) +
  geom_point() +
  geom_smooth(method = "lm")

d0 %>% 
  ggplot(aes(ct_last_time, ctc_prop)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### CT and AW rate variables and symptoms
```{r}
cor.test(d0$ctc_rate, d0$itsea_symptoms)
cor.test(d0$awc_rate, d0$itsea_symptoms)
```

### Covary duration of recording

#### AW
```{r}
sym_aw_duration <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(awc_prop) + 
      scale(cesd_t4) +
      scale(NEG) +
      scale(duration_analyzed),
    data = d0 
  )

calc.relimp(sym_aw_duration)
model_parameters(sym_aw_duration)
```
```{r}
sym_ct_duration <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(ctc_prop) + 
      scale(cesd_t4) +
      scale(NEG) +
      scale(duration_analyzed),
    data = d0 
  )

calc.relimp(sym_ct_duration)
model_parameters(sym_ct_duration)
```

### SES and LENA measures
```{r}
# education
cor.test(d0$ctc_hour_max, d0$education, method = "spearman")
cor.test(d0$awc_hour_max, d0$education, method = "spearman")

cor.test(d0$ctc_prop, d0$education, method = "spearman")
cor.test(d0$awc_prop, d0$education, method = "spearman")

# income (HUD ratio)
cor.test(d0$ctc_hour_max, d0$income_needs)
cor.test(d0$awc_hour_max, d0$income_needs)

cor.test(d0$ctc_prop, d0$income_needs)
cor.test(d0$awc_prop, d0$income_needs)

# income (FPL ratio)
cor.test(d0$ctc_hour_max, d0$inr_fpl)
cor.test(d0$awc_hour_max, d0$inr_fpl)

cor.test(d0$ctc_prop, d0$inr_fpl)
cor.test(d0$awc_prop, d0$inr_fpl)
```

### Covary for SES

#### AW 
```{r}
sym_aw_ses <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(awc_prop) + 
      scale(cesd_t4) +
      scale(NEG) +
      scale(income_needs) +
      scale(education),
    data = d0 
  )

calc.relimp(sym_aw_ses)
model_parameters(sym_aw_ses)
```

#### CT
```{r}
sym_ct_ses <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(ctc_prop) +
      scale(cesd_t4) +
      scale(NEG) +
      scale(income_needs) +
      scale(education),
    data = d0 
  )

calc.relimp(sym_ct_ses)
model_parameters(sym_ct_ses)
```
## Export data
```{r}
d0_share <-
  d0 %>% 
  dplyr::select(
    ID,
    male,
    age_lena_d1,
    age_18mo,
    mom_age,
    education,
    education_txt,
    annual_income_txt,
    income_needs,
    inr_fpl,
    mom_race,
    white,
    mom_latinx,
    secondlang,
    day_type,
    NEG,
    cesd_t3,
    cesd_t4,
    duration_total,
    duration_analyzed,
    ct_first_time,
    ct_last_time,
    percent_mother,
    ctc_prop,
    awc_prop,
    ctc_hour_max,
    awc_hour_max,
    cvc_hour_max,
    ctc_rate,
    awc_rate,
    sens_sfp,
    sens_fp,
    itsea_symptoms,
    itsea_concern_ied
  )
```

```{r}
var_label(d0_share) <- 
  list(
    male = "Infant sex",
    age_lena_d1 = "Infant age at LENA assessment",
    age_18mo = "Child age at toddler assessment",
    mom_age = "Maternal age at infant assessment",
    education = "Maternal education (ordinal)",
    education_txt = "Maternal education (labeled)",
    annual_income_txt = "Household annual income reported in bins",
    income_needs = "Income-to-needs ratio based on 2017 HUD limits for SC county",
    inr_fpl = "Income-to-needs ratio based on 2017 FPL",
    mom_race = "Maternal self-reported race",
    mom_latinx = "Maternal self-reported ethnicity",
    secondlang = "Language additional to English spoken at home",
    day_type = "Weekend or weekday for LENA recording",
    NEG = "Infant negative affectivity on IBQ-R-SF",
    cesd_t3 = "Maternal depressive symptoms at infant assessment",
    cesd_t4 = "Maternal depressive symptoms at toddler assessment",
    duration_total = "Total duration of LENA recording",
    duration_analyzed = "Duration of portion of LENA recording analyzed",
    ct_first_time = "Time of first CT of the recording day",
    ct_last_time = "Time of last CT of the recording day",
    percent_mother = "Percent of recording day with mother",
    ctc_prop = "Consistency of conversational turns",
    awc_prop = "Consistency of adult words",
    ctc_hour_max = "Quantity of conversational turns",
    awc_hour_max = "Quantity of adult words",
    cvc_hour_max = "Quantity of infant vocalizations",
    ctc_rate = "Hourly rate of conversational turns",
    awc_rate = "Hourly rate of adult words",
    sens_sfp = "Maternal sensitivity to distress",
    sens_fp = "Maternal sensitivity during play",
    itsea_symptoms = "Total symptoms on abbreviated ITSEA",
    itsea_concern_ied = "Met cutoff for clinical concern for at least one symptom subscale"
  )

metadata(d0_share)$name <- "Analysed data for \"Dimensions of the language environment in infancy and symptoms of psychopathology in toddlerhood\" (King et al. 2020)"
```

```{r}
#write_csv(d0_share, "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science#/sub3/king_analyzed_data_20201220.csv")
```

