---
title: "Association of linguistic input at age 6 months and symptoms of psychopathology at age 18 months"
author: "Lucy King"
output: html_document
---

# Environment
```{r setup, include=FALSE}
# Libraries
library(tidyverse)
library(lubridate)
library(corrr)
library(corrplot)
library(ggpubr)
library(modelr)
library(effectsize)
library(glmpath)
library(glmnet)
library(MASS)
library(ggcorrplot)
library(broom)
library(ggfortify)
library(relaimpo)
library(olsrr)

# Files 
home <- "~/Box/Mooddata_Coordinating/BABIES/Data/final_scored_data/"

eligible_id_file <- "~/Box/lucy_king_files/BABIES/lena_symptoms/inclusion_tracker.xlsx"
fp_intervention_id <- paste0(home, "lab_caregiving_behavior/free_play_intervention_assignment.csv")
itsea_file <- paste0(home, "survey_18month/itsea_scored_20200414.csv")
demographics_file <- paste0(home, "demographics/demo_6mo_cleaned_final.csv")
lena_file <- paste0(home, "LENA/lena_final_wf_day1_20200208_upto9months.csv")
ibq_file <- paste0(home, "IBQ/ibq_scored_final.csv")
sfp_care_file <- paste0(home, "lab_caregiving_behavior/PCIRS_sfp_wf_complete.csv")
fp_care_file <- paste0(home, "lab_caregiving_behavior/free_play_wf_8min_20200507.csv")
cesd_file <- paste0(home, "CESD/cesd_wf_20200507.csv")

# Functions
source("identify_outliers_histogram.R")
source("winsorize.R")

theme_lena <-
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  )

# Display options
options(scipen = 999)

```


# Read in and join data
```{r}
d0 <-
  readxl::read_xlsx(eligible_id_file, sheet = "data_filtered_eligible") %>% 
  dplyr::select(
    ID,
    eligible,
    included
  ) %>% 
  left_join(
    read_csv(demographics_file) %>% 
      dplyr::select(
        ID,
        baby_dob,
        education,
        education_txt,
        annual_income_txt,
        income_needs,
        male,
        mom_age,
        mom_latinx,
        mom_race,
        age_behav,
        secondlang,
        secondlang_type
      ),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(itsea_file) %>% 
      dplyr::select(
        ID,
        survey_date,
        itsea_symptoms,
        itsea_activity,
        itsea_agress,
        itsea_gad,
        itsea_negemo,
        itsea_depp,
        itsea_intern,
        itsea_extern,
        # itsea symptoms items
        itsea_a2, 
        itsea_a4, 
        itsea_b8, 
        itsea_b10, 
        itsea_b27, 
        itsea_b32,
        itsea_a28, 
        itsea_a30, 
        itsea_a33, 
        itsea_b3, 
        itsea_b9, 
        itsea_b16, 
        itsea_b30, 
        itsea_b34, 
        itsea_b44, 
        itsea_b88, 
        itsea_c2, 
        itsea_e4,
        itsea_a9, 
        itsea_a39, 
        itsea_a40, 
        itsea_b43, 
        itsea_b76, 
        itsea_b81, 
        itsea_b84, 
        itsea_b91, 
        itsea_b92,
        itsea_a3, 
        itsea_a12, 
        itsea_a35, 
        itsea_b37, 
        itsea_b70, 
        itsea_b86, 
        itsea_e5, 
        itsea_e6, 
        itsea_e11,
        itsea_a7, 
        itsea_a21, 
        itsea_a23, 
        itsea_b31, 
        itsea_b45, 
        itsea_b50, 
        itsea_b53, 
        itsea_b59, 
        itsea_b65, 
        itsea_b66, 
        itsea_b74, 
        itsea_b80, 
        itsea_b85,
        itsea_concern_ied
      ), 
    by = "ID"
  ) %>% 
  mutate(
    survey_date =  parse_date_time(survey_date, orders = c("mdy HM")),
    age_18mo = (baby_dob %--% survey_date) / months(1)
  ) %>% 
  left_join(
    read_csv(fp_care_file) %>% 
      dplyr::select(ID, negmood_FP, sens_FP),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(sfp_care_file) %>% 
      dplyr::select(ID, sens_R_M, sens_M),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(ibq_file) %>% 
      dplyr::select(
        ID,
        #IBQ NEG items
        ibq_64, ibq_74, ibq_75, ibq_32, ibq_79, ibq_80,
        ibq_2, ibq_3_r, ibq_4, ibq_21, ibq_52, ibq_53, ibq_62,
        ibq_22, ibq_76, ibq_77, ibq_78, ibq_87, ibq_89,
        ibq_36, ibq_37, ibq_38, ibq_63, ibq_71, ibq_72,
        NEG
      ),
    by = "ID"
  ) %>% 
  left_join(read_csv(lena_file), by = "ID") %>% 
  left_join(read_csv(cesd_file), by = "ID") %>% 
  left_join(read_csv(fp_intervention_id), by = "ID") %>% 
  rename(
    age_lena_d1 = age,
    sens_r_sfp = sens_R_M,
    sens_sfp = sens_M,
    sens_fp = sens_FP,
    negmood_fp = negmood_FP
  ) 
```

# Retention / Missing data

Dyads who completed the 6-month assessment (either lab or home observation), were eligible for the 18-month assesment, and completed the 18-month assessment are included in the current analyses. 

```{r}
#filter to only include those with 18-month symptoms
d0_d1 <-
  d0 %>% 
  filter(
    !is.na(itsea_symptoms)
  ) 

##missing LENA
d0_d1 %>% 
  count(is.na(ctc_hour_max))

# filter to only include those with LENA
d0_d1 <-
  d0_d1 %>% 
  filter(!is.na(ctc_hour_max))

#missing SFP
d0_d1 %>% 
  count(is.na(sens_sfp))

#missing FP
d0_d1 %>% 
  count(is.na(sens_fp))

#missing IBQ
d0_d1 %>% 
  count(is.na(NEG))
```

# Descriptive statistics 

```{r}
### dyads in FP intervention
d0_d1 %>% 
  count(group)
```


## IBQ, ITSEA, CESD reliability

```{r}
d0_d1 %>% 
  dplyr::select(ibq_64:ibq_72) %>% 
  psych::alpha(check.keys = TRUE)
```

```{r}
d0_d1 %>% 
  dplyr::select(itsea_a2:itsea_b85, -itsea_c2) %>% 
  psych::alpha()
```

```{r}
d0_d1 %>% 
  summarize_at(
    vars(
      age_18mo,
      age_lena_d1,
      NEG,
      duration,
      cvc_hour_max,
      ctc_hour_max,
      awc_hour_max,
      ctc_prop,
      awc_prop,
      sens_r_sfp,
      sens_fp,
      negmood_fp,
      itsea_symptoms,
      mom_age,
      income_needs,
      percent_mother,
      cesd_t3,
      cesd_t4
    ), 
    funs(mean, sd, min, max), na.rm = TRUE
  )
```
Race:
1, American Indian or Alaska Native
2, Asian
3, Black or African American
4, Native Hawaiian or Other Pacific Islander
5, White
6, Other

Education:
0, No schooling completed
2, Nursery school to 8th grade
3, Some high school, no diploma
4, High school graduate, diploma or the equivalent (i.e. GED)
5, Some college credit, no degree
6, Trade/technical/vocational training
7, Associate degree
8, Bachelor's degree
9, Graduate degree
10, Other

Income:
1, $0-5,000
2, $5,001-15,000
3, $15,001-30,000
4, $30,001-60,000
5, $60,001-90,000
6, $90,001-150,000
7, Greater than $150,000
```{r}
d0_d1 %>% 
  count(male) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1 %>% 
  filter(!is.na(duration)) %>% 
  count(duration >= 12) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1 %>% 
  count(itsea_concern_ied) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1 %>% 
  count(income_needs < 1) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1 %>% 
  count(mom_latinx) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1 %>% 
  count(mom_race) %>% 
  mutate(
    per = n / sum(n)
  ) %>% 
  arrange(desc(n))

d0_d1 %>% 
  count(education_txt) %>% 
  mutate(
    per = n / sum(n)
  ) %>% 
  arrange(education_txt)

d0_d1 %>% 
  count(day_type) %>% 
  mutate(per = n / sum(n))

d0_d1 %>% 
  count(cesd_t3 >= 16) %>% 
  mutate(per = n / sum(n))

d0_d1 %>% 
  count(cesd_t4 >= 16) %>% 
  mutate(per = n / sum(n))

```


## Distributions 

```{r}
d0_d1 %>% 
  identify_outliers_hist(x = cvc_hour_max)

d0_d1 %>% 
  identify_outliers_hist(x = ctc_hour_max)

d0_d1 %>% 
  identify_outliers_hist(x = ctc_prop)

d0_d1 %>% 
  identify_outliers_hist(x = awc_hour_max)

d0_d1 %>% 
  identify_outliers_hist(x = awc_prop)

d0_d1 %>% 
  identify_outliers_hist(x = sens_r_sfp)

d0_d1 %>% 
  identify_outliers_hist(x = sens_fp)

d0_d1 %>% 
  identify_outliers_hist(x = itsea_symptoms)

d0_d1 %>% 
  identify_outliers_hist(x = cesd_t3)

d0_d1 %>% 
  identify_outliers_hist(x = cesd_t4)


d0_d1 %>% 
  ggplot(aes(percent_mother)) +
  geom_histogram() +
  theme_lena +
  labs(
    x = "% of the LENA recording day infant with mother"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/percent_mom_hist.png",
  width = 7,
  height = 5
)
```


## Correlations

``` {r}
corr_lena_plot <- 
  d0_d1 %>%
  dplyr::select(
    `AW consistency` = awc_prop,
    `CT consistency` = ctc_prop,
    `AW quantity` = awc_hour_max,
    `CT quantity` = ctc_hour_max,
    `Infant vocalizations` = cvc_hour_max
  ) %>% 
  cor(use = "complete.obs", method = "pearson")

ggcorrplot(
  corr_lena_plot, 
  hc.order = FALSE, 
  type = "upper", 
  lab = TRUE,
  ggtheme = theme_void,
  legend.title = "Pearson\ncorrelation coefficient",
  digits = 2
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/lena_corr_plot.png", 
  height = 5, 
  width = 5
)
```

```{r}
cor.test(d0_d1$cesd_t3, d0_d1$awc_hour_max)
cor.test(d0_d1$cesd_t3, d0_d1$awc_prop)
cor.test(d0_d1$cesd_t3, d0_d1$ctc_hour_max)
cor.test(d0_d1$cesd_t3, d0_d1$ctc_prop)
cor.test(d0_d1$cesd_t3, d0_d1$cvc_hour_max)
cor.test(d0_d1$cesd_t3, d0_d1$sens_r_sfp)
cor.test(d0_d1$cesd_t3, d0_d1$sens_fp)

cor.test(d0_d1$NEG, d0_d1$awc_hour_max)
cor.test(d0_d1$NEG, d0_d1$awc_prop)
cor.test(d0_d1$NEG, d0_d1$ctc_hour_max)
cor.test(d0_d1$NEG, d0_d1$ctc_prop)
cor.test(d0_d1$NEG, d0_d1$cvc_hour_max)
cor.test(d0_d1$NEG, d0_d1$sens_r_sfp)
cor.test(d0_d1$NEG, d0_d1$sens_fp)
```

```{r}
cor.test(d0_d1$awc_hour_max, d0_d1$sens_fp)
cor.test(d0_d1$awc_hour_max, d0_d1$sens_r_sfp)

cor.test(d0_d1$ctc_hour_max, d0_d1$sens_fp)
cor.test(d0_d1$ctc_hour_max, d0_d1$sens_r_sfp)

cor.test(d0_d1$awc_prop, d0_d1$sens_fp)
cor.test(d0_d1$awc_prop, d0_d1$sens_r_sfp)

cor.test(d0_d1$ctc_prop, d0_d1$sens_fp)
cor.test(d0_d1$ctc_prop, d0_d1$sens_r_sfp)

cor.test(d0_d1$cvc_hour_max, d0_d1$sens_fp)
cor.test(d0_d1$cvc_hour_max, d0_d1$sens_r_sfp)

cor.test(d0_d1$sens_fp, d0_d1$sens_r_sfp)
```

## t-tests by bilingual status
```{r}
t.test(d0_d1$awc_hour_max ~ d0_d1$secondlang)

t.test(d0_d1$ctc_hour_max ~ d0_d1$secondlang)

t.test(d0_d1$ctc_prop ~ d0_d1$secondlang)

t.test(d0_d1$awc_prop ~ d0_d1$secondlang)

t.test(d0_d1$itsea_symptoms ~ d0_d1$secondlang)
```

## t-tests by day type
```{r}
t.test(d0_d1$awc_hour_max ~ d0_d1$day_type)

t.test(d0_d1$ctc_hour_max ~ d0_d1$day_type)

t.test(d0_d1$ctc_prop ~ d0_d1$day_type)

t.test(d0_d1$awc_prop ~ d0_d1$day_type)
```


## Histograms
```{r}
d0_d1 %>% 
  dplyr::select(
    ctc_hour_max,
    awc_hour_max,
    ctc_prop,
    awc_prop,
    sens_fp,
    sens_r_sfp
  ) %>% 
  gather(measure, value, ctc_hour_max:sens_r_sfp) %>% 
  mutate(
    measure = factor(
      measure,
      levels = c(
        "awc_prop",
        "ctc_prop",
        "awc_hour_max",
        "ctc_hour_max",
        "sens_r_sfp",
        "sens_fp"
      ),
      labels = c(
        "AW consistency", 
        "CT consistency", 
        "AW quantity", 
        "CT quantity",
        "Sensitivity to distress",
        "Sensivity during play"
        )
    )
  ) %>% 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(.~measure, scales = "free", ncol = 2, nrow = 3) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 18),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Caregiving input value"
  )


ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/care_hist_ggpanels.png",
  width = 7.5,
  height = 8
)
```

```{r}
d0_d1 %>% 
  ggplot(aes(itsea_symptoms)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq.int(0, 3, .25)) +
  theme_lena +
  labs(
    x = "Toddler symptoms of psychopathology"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/itsea_hist.png"
)
```

```{r}
d0_d1 %>% 
  dplyr::select(
    Activity = itsea_activity,
    Aggression = itsea_agress,
    Depression = itsea_depp,
    Anxiety = itsea_gad,
    `Negative emotion` = itsea_negemo
  ) %>% 
  gather(itsea_subscale, score, Activity:`Negative emotion`) %>% 
  ggplot(aes(score)) +
  geom_histogram(bins = 12) +
  theme_lena +
  theme(
    strip.text = element_text(size = 16)
  ) +
  facet_wrap(.~itsea_subscale, scales = "free") +
  labs(
    x = "Subscale score"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/itsea_subscale_hist.png",
  width = 10,
  height = 7
)
```

```{r}
d0_d1 %>% 
  ggplot(aes(cesd_t3)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq.int(0, 50, 5)) +
  theme_lena +
  labs(
    x = "Maternal depressive symptoms\n at infant age 6 months"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/cesd6_hist.png"
)
```

```{r}
d0_d1 %>% 
  ggplot(aes(cesd_t4)) +
  geom_histogram() +
  scale_x_continuous(breaks = seq.int(0, 50, 5)) +
  theme_lena +
  labs(
    x = "Maternal depressive symptoms\n at infant age 18 months"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/cesd18_hist.png"
)
```

# Results 

## Associations of LENA measures with ITSEA symptoms 

### Effect code categorical variables

```{r}
d0_d1 <-
  d0_d1 %>% 
  mutate(
    white = as.factor(
      if_else(
        mom_race == "White", 
        1, 0
      )
    ),
    male = as.factor(male),
    mom_latinx = as.factor(mom_latinx)
  )

contrasts(d0_d1$male) <- c(-.5, .5)
contrasts(d0_d1$white) <- c(-.5, .5)
contrasts(d0_d1$mom_latinx) <- c(-.5, .5)
```

## Step-wise regression using stepAIC (MASS package) with LENA variables and covariates

### Fit initial linear regression
```{r}
d0_d1_step <-
  d0_d1 %>% 
  dplyr::select(
    ID, 
    itsea_symptoms,
    ctc_prop,
    awc_prop,
    ctc_hour_max,
    awc_hour_max,
    cvc_hour_max,
    cesd_t4,
    cesd_t3,
    NEG,
    age_lena_d1,
    mom_age,
    income_needs,
    education,
    white,
    male,
    mom_latinx
  ) %>% 
  na.omit() %>% 
  rowid_to_column()
```

```{r}
sym_lena <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(ctc_prop) +
      scale(awc_prop)  +
      scale(ctc_hour_max) +
      scale(awc_hour_max) +
      scale(cvc_hour_max) +
      scale(cesd_t4) +
      scale(cesd_t3) +
      scale(NEG) +
      scale(age_lena_d1) +
      scale(mom_age) +
      scale(income_needs) +
      scale(education) +
      white +
      male +
      mom_latinx,
    data = d0_d1_step 
  )
summary(sym_lena)
```

### Model diagnostics
http://www.sthda.com/english/articles/39-regression-model-diagnostics/161-linear-regression-assumptions-and-diagnostics-in-r-essentials/

NOTE that numbers represent the row index number
```{r}
sym_lena.diag.metrics <- 
  augment(sym_lena) %>% 
  rowid_to_column()
```

When data points have high Cook’s distance scores and are to the upper or lower right of the leverage plot, they have leverage meaning they are influential to the regression results. The regression results will be altered if we exclude those cases. I have no such cases. 
```{r}
#outliers and high leverage
plot(sym_lena, 5)
plot(sym_lena, 4)
```

https://cran.r-project.org/web/packages/olsrr/vignettes/regression_diagnostics.html

Variance inflation factors measure the inflation in the variances of the parameter estimates due to collinearities that exist among the predictors. It is a measure of how much the variance of the estimated regression coefficient βk is “inflated” by the existence of correlation among the predictor variables in the model. A VIF of 1 means that there is no correlation among the kth predictor and the remaining predictor variables, and hence the variance of βk is not inflated at all. The general rule of thumb is that VIFs exceeding 4 warrant further investigation, while VIFs exceeding 10 are signs of serious multicollinearity requiring correction.

```{r}
#collinearity
ols_vif_tol(sym_lena)
```

### Fit stepwise regression (complete data)
```{r}
symp_lena_step <- stepAIC(sym_lena, direction = "both")
summary(symp_lena_step)
calc.relimp(symp_lena_step)
confint(symp_lena_step)
```

```{r}
confint_lena_step <- 
  confint(symp_lena_step) %>% 
  as.tibble() 

tidy_symp_lena_step <- 
  tidy(symp_lena_step) %>% 
  rename(
    Variable = term,
    B = estimate,
    SE = std.error,
    t = statistic,
    p = p.value
  ) %>% 
  mutate(
    Variable = recode(
      Variable,
      "(Intercept)" = "(Intercept)",
      "scale(awc_prop, scale = FALSE)" = "AW consistency",
      "scale(cvc_hour_max, scale = FALSE)" = "Infant vocalizations",
      "scale(cesd_t4, scale = FALSE)" = "Maternal depressive symptoms",
      "scale(NEG, scale = FALSE)" = "Infant negative affectivity"  
    )
  ) %>% 
  cbind(confint_lena_step)  %>% 
  mutate_at(
    vars(B:t, `2.5 %`, `97.5 %`),
    round, 2
  ) %>% 
  mutate_at(
    vars(p),
    round, 3
  ) %>% 
  unite(col = "95% CI", `2.5 %`, `97.5 %`, sep = ",") %>% 
  dplyr::select(
    Variable, 
    B, 
    everything()
  ) 

write.table(
  tidy_symp_lena_step, "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/stepwise_table",
  row.names = FALSE,
  quote = FALSE,
  sep = ","
)
```

### Fit stepwise regression (excluding higher leverage cases)
```{r}
d0_d1_step_diag <-
  d0_d1 %>% 
  filter(ID != 50, ID != 1027, ID != 1037) %>% 
  dplyr::select(
    ID, 
    itsea_symptoms,
    ctc_prop,
    awc_prop,
    ctc_hour_max,
    awc_hour_max,
    cvc_hour_max,
    cesd_t4,
    cesd_t3,
    NEG,
    age_lena_d1,
    mom_age,
    income_needs,
    education,
    white,
    male,
    mom_latinx
  ) %>% 
  na.omit()
```

```{r}
sym_lena_diag <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(ctc_prop) +
      scale(awc_prop)  +
      scale(ctc_hour_max) +
      scale(awc_hour_max) +
      scale(cvc_hour_max) +
      scale(cesd_t4) +
      scale(cesd_t3) +
      scale(NEG) +
      scale(age_lena_d1) +
      scale(mom_age) +
      scale(income_needs) +
      scale(education) +
      white +
      male +
      mom_latinx,
    data = d0_d1_step_diag 
  )
summary(sym_lena_diag)
```

```{r}
sym_lena_diag_step <- stepAIC(sym_lena_diag, direction = "both")
summary(sym_lena_diag_step)
calc.relimp(sym_lena_diag_step)
confint(sym_lena_diag_step)
```

### Create residualized symptoms variable for covariates in best fitting model
```{r}
best_cov <- lm(
  itsea_symptoms ~
    scale(cesd_t4, scale = FALSE) +
    scale(NEG, scale = FALSE),
  data = d0_d1
)

d0_d1 <-
  d0_d1 %>% 
  add_residuals(best_cov, var = "resid_symptoms")
```

### Plot associations 

### AW consistency
```{r}
d0_d1 %>% 
  ggplot(aes(awc_prop, resid_symptoms)) +
  geom_point(size = 4, alpha = 1/2) +
  geom_smooth(method = "lm", size = 3, color = "black") +
  scale_x_continuous(
    breaks = seq.int(0, 1, .10),
    labels = scales::percent_format(accuracy = 1)
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 17),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Consistency of adult words in daily life",
    y = "Toddler symptoms of psychopathology\n(residuals)"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/aw_prop_symptoms.png",
  width = 7,
  height = 6
)
```

### Infant vocalizations
```{r}
d0_d1 %>% 
  ggplot(aes(cvc_hour_max, resid_symptoms)) +
  geom_point(size = 4, alpha = 1/2) +
  geom_smooth(method = "lm", size = 3, color = "black") +
  scale_x_continuous(
    breaks = seq.int(0, 1000, 100)
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 17),
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Quantity of infant vocalizations in daily life",
    y = "Toddler symptoms of psychopathology\n(residuals)"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/plots/cv_max_symptoms.png",
  width = 7,
  height = 6
)
```

## Elastic net regression with LENA variables and covariates

### Prepare data
```{r}
z_score <- function(x) {
  diff_mu <- x - mean(x, na.rm = T)
  sd <- sd(x, na.rm = T)
  diff_mu / sd
}
```

```{r}
d0_en <-
  d0_d1 %>% 
  dplyr::select(
    itsea_symptoms,
    ctc_prop,
    awc_prop,
    ctc_hour_max,
    awc_hour_max,
    cvc_hour_max,
    cesd_t4,
    cesd_t3,
    NEG,
    age_lena_d1,
    mom_age,
    income_needs,
    education,
    white,
    male,
    mom_latinx
  ) %>% 
  mutate_at(
    vars(itsea_symptoms:education),
    funs(z_score)
  ) %>% 
  mutate_at(
    vars(white:mom_latinx),
    funs(as.numeric)
  ) %>% 
  mutate_at(
    vars(white, male, mom_latinx),
    funs(. %>% dplyr::recode("2" = 1, "1" = 0))
  ) %>% 
  na.omit()
```

```{r}
predictors <- 
  d0_en %>% 
  dplyr::select(-itsea_symptoms) %>% 
  as.matrix()
  
symptoms <- z_score(d0_en$itsea_symptoms)
```

### Run model over 100 increasingly relaxed lambda values
```{r}
fit_net <- glmnet(predictors, symptoms, family = "gaussian")
plot(fit_net, label = TRUE)
```

### Run leave-one-out cross validation to identify optimal lambda value
If nfolds is set to the sample size = leave-one-out CV
https://sciphy-stats.com/post/2019-01-25-finalizing-glmnet-models/
Running the CV 100 times to minimize simulation error. 
```{r}
# run cross validation
set.seed(123)

lambdaMins <- c()

for (i in 1:100) {
  fit_cv <- cv.glmnet(
    predictors, 
    symptoms, 
    type.measure = "mse", 
    family = "gaussian", 
    alpha = .5,
    nfolds = 91,
    grouped = FALSE
  )
  lambdaMins <- cbind(lambdaMins, fit_cv$lambda.min)
}
```

```{r}
# save lambda corresponding to minimal mse to a tibble
lambdaMins <- as_tibble(lambdaMins)
lambdaMins <-
  lambdaMins %>%
  gather(run, lambdaMin, V1:V100) 

# calculate mean lambda value 
lambdaFinal <- mean(lambdaMins$lambdaMin)
```

### Print coefficients for model with optimal lambda 
```{r}
fit_net_final <- glmnet(predictors, symptoms, family = "gaussian", alpha = .5, lambda = lambdaFinal)
coefs <- coef(fit_net_final) 

en_coef_tbl <- 
  tidy(coef(fit_net_final))  %>% 
  dplyr::select(-column) %>% 
  rename(
    Variable = row,
    `Elastic net estimate` = value
  ) %>% 
  mutate(
    Variable = recode(
      Variable,
      "(Intercept)" = "(Intercept)",
      "ctc_prop" = "CT consistency",
      "awc_prop" = "AW consistency",
      "cvc_hour_max" = "Infant vocalizations",
      "cesd_t4" = "Maternal depressive symptoms",
      "NEG" = "Infant negative affectivity"  ,
      "age_lena_d1" = "Infant age",
      "mom_age" = "Maternal age",
      "mom_latinx" = "Latinx ethnicity"
    )
  ) %>% 
  mutate_at(vars(`Elastic net estimate`) , round, 2)
```

### Calculate EN model performance
```{r}
# save model matrix
X <- model.matrix(itsea_symptoms ~ ., d0_en)

# get predicted values
y_hat <- matrix(X, ncol = 16) %*% as.matrix(coefs)

# calculate R-squared
r_sq <- 1 - (sum((y_hat[, 1] - symptoms)^2) / sum((symptoms - mean(symptoms))^2))
```

### Evaluate model performance

```{r}
for (i in 1000) {
  fit_rr_cv <- cv.glmnet(predictors, symptoms, type.measure = "mse", alpha = .5, family = "gaussian")
  lambdaMins <- cbind(lambdaMins, fit_rr_cv$lambda.min)
}
```

## Additional analyses

### Covarying maternal sensitivity in best fitting model
```{r}
sym_lena_ms <- 
  lm(
    scale(itsea_symptoms) ~ 
      scale(awc_prop) + 
      scale(cvc_hour_max) +
      scale(cesd_t4) +
      scale(NEG) +
      scale(sens_fp) +
      scale(sens_r_sfp),
    data = d0_d1 
  )

summary(sym_lena_ms)
calc.relimp(sym_lena_ms)
confint(sym_lena_ms)
```

## Who is talking?

```{r}
d0_d1_follow <-
  d0_d1 %>% 
  filter(
    lena_caregivers___1 != 0 |
      lena_caregivers___2 != 0 |
      lena_caregivers___3 != 0 |
      lena_caregivers___4 != 0 |
      lena_caregivers___5 != 0 |
      lena_caregivers___6 != 0 |
      lena_caregivers___7 != 0 |
      lena_caregivers___8 != 0 |
      lena_caregivers___9 != 0 |
      lena_caregivers___10 != 0
  ) %>% 
  rename(
    mom_present = lena_caregivers___1,
    dad_present = lena_caregivers___2,
    daycare_present = lena_caregivers___3,
    nanny_present = lena_caregivers___4,
    sibling_present = lena_caregivers___5,
    cousin_present = lena_caregivers___6,
    grandparent_present = lena_caregivers___7,
    friends_present = lena_caregivers___8,
    relative_present = lena_caregivers___9, #any other relative
    other_present = lena_caregivers___10
  )
```

```{r}
d0_d1_follow %>% 
  count(!is.na(percent_mother))

d0_d1_follow %>%
  summarise_at(
    vars(percent_mother),
    funs(mean, median, sd, min, max), na.rm = TRUE
  )

cor.test(d0_d1_follow$percent_mother, d0_d1_follow$awc_prop, method = "spearman")
cor.test(d0_d1_follow$percent_mother, d0_d1_follow$cvc_hour_max, method = "spearman")
```

```{r}
d0_d1_follow %>%
  count(mom_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(dad_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(mom_present, dad_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(daycare_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(nanny_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(grandparent_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(friends_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(relative_present) %>% 
  mutate(
    per = n / sum(n)
  )

d0_d1_follow %>%
  count(
    mom_present, 
    dad_present, 
    daycare_present, 
    nanny_present, 
    grandparent_present
  ) %>% 
  mutate(
    per = n / sum(n)
  ) %>% 
  arrange(n)

```

```{r}
d0_d1_follow %>% 
  count(percent_mother)
```


```{r}
d0_d1_follow <-
  d0_d1_follow %>% 
  mutate(
    n_caregivers = 
      mom_present +
      dad_present +
      daycare_present +
      grandparent_present +
      nanny_present +
      friends_present
  )

t.test(d0_d1_follow$ctc_prop ~ d0_d1_follow$dad_present)
t.test(d0_d1_follow$ctc_prop ~ d0_d1_follow$grandparent_present)
```


```{r}
#write_csv(d0_d1, "~/Box/lucy_king_files/BABIES/lena_symptoms/manuscript/dev_science/king_analyzed_data_20200528.csv")
```

