---
title: "Score ITSEA"
output: html_notebook
---

```{r}
# Libraries
library(tidyverse)
library(lubridate)
library(purrr)
library(psych)
library(GPArotation)

# Files

q_18month_cs_file <- "~/Desktop/BABIES/lena_symptoms/data_final/qs_18mo_cs_datecleaned_20190726.csv"
q_18month_long_file <- "~/Desktop/BABIES/lena_symptoms/data_final/qs_18mo_long_datecleaned_20190726.csv"
demographics_file <- "~/Desktop/BABIES/lena_symptoms/data_final/demographics_20190726.csv"

theme_lena <-
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  )
```

# Read in data
```{r}
itsea <-
  read_csv(q_18month_cs_file) %>% 
  rename(ID = record_id) %>% 
  mutate(ID = as.integer(ID)) %>% 
  select(-redcap_event_name:-questionnaires_18mo_timestamp) %>% 
  bind_rows(
    read_csv(q_18month_long_file) %>% 
      rename(ID = record_id) %>% 
      mutate(ID = as.integer(ID)) %>%
      select(-redcap_event_name:-month_survey_timestamp) 
  ) %>% 
  select(ID, survey_date, itsea_a2:itsea_e11) %>% 
  left_join(
    read_csv(demographics_file) %>% 
      select(ID, male, baby_dob),
    by = "ID"
  )
```

# Score ITSEA

## Compute number of non-missing items for each subscale
```{r}
itsea <-
  itsea %>% 
  group_by(ID) %>% 
  mutate(
    n_itsea_activity = sum(
      !is.na(
        c(
          itsea_a2, 
          itsea_a4, 
          itsea_b8, 
          itsea_b10, 
          itsea_b27, 
          itsea_b32
        )
      )
    ),
    n_itsea_agress = sum(
      !is.na(
        c(
          itsea_a28, 
          itsea_a30, 
          itsea_a33, 
          itsea_b3, 
          itsea_b9, 
          itsea_b16, 
          itsea_b30, 
          itsea_b34, 
          itsea_b44, 
          itsea_b88, 
          itsea_c2, 
          itsea_e4
        )
      )
    ),
    n_itsea_depp = sum(
      !is.na(
        c(
          itsea_a9, 
          itsea_a39, 
          itsea_a40, 
          itsea_b43, 
          itsea_b76, 
          itsea_b81, 
          itsea_b84, 
          itsea_b91, 
          itsea_b92
        )
      )
    ),
    n_itsea_empath = sum(
      !is.na(
        c(
          itsea_a22, 
          itsea_b51, 
          itsea_b56, 
          itsea_b57, 
          itsea_b63, 
          itsea_b73, 
          itsea_c4
        )
      )
    ),
    n_itsea_gad = sum(
      !is.na(
        c(
          itsea_a3, 
          itsea_a12, 
          itsea_a35, 
          itsea_b37, 
          itsea_b70, 
          itsea_b86, 
          itsea_e5, 
          itsea_e6, 
          itsea_e11
        )
      )
    ),
    n_itsea_play = sum(
      !is.na(
        c(
          itsea_a25, 
          itsea_a31, 
          itsea_b25, 
          itsea_b38, 
          itsea_b67, 
          itsea_b71
        )
      )
    ),
    n_itsea_negemo = sum(
      !is.na(
        c(
          itsea_a7, 
          itsea_a21, 
          itsea_a23, 
          itsea_b31, 
          itsea_b45, 
          itsea_b50, 
          itsea_b53, 
          itsea_b59, 
          itsea_b65, 
          itsea_b66, 
          itsea_b74, 
          itsea_b80, 
          itsea_b85
        )
      )
    ),
    n_itsea_social = sum(
      !is.na(
        c(
          itsea_a10, 
          itsea_a13, 
          itsea_a15, 
          itsea_b6, 
          itsea_b19, 
          itsea_b26, 
          itsea_b28, 
          itsea_b42, 
          itsea_b60, 
          itsea_b90
        )
      )
    )
  ) %>% 
  ungroup()
```


## Calculate subscale scores

Using scoring specific by ITSEA manual. For each subscale, calculate the mean by taking the sum and dividing by the number of items the both responded to. If the mother missed too many items, replace subscale score with NA.
```{r}
itsea <-
  itsea %>% 
  #reverse code items
  mutate_at(
    vars(
      itsea_b43, itsea_b65
    ),
    funs(
      . %>%
        dplyr::recode(
          "0" = 2, 
          "1" = 1, 
          "2" = 0
        )
    )
  ) %>% 
  group_by(ID) %>% 
  mutate(
    itsea_activity = sum(
      c(
        itsea_a2, 
        itsea_a4, 
        itsea_b8, 
        itsea_b10, 
        itsea_b27, 
        itsea_b32
      ), na.rm = TRUE
    ) / n_itsea_activity,
    itsea_activity = if_else(
      n_itsea_activity < 5, 
      NA_real_, itsea_activity
    ),
    itsea_agress = sum(
      c(
        itsea_a28, 
        itsea_a30, 
        itsea_a33, 
        itsea_b3, 
        itsea_b9, 
        itsea_b16, 
        itsea_b30, 
        itsea_b34, 
        itsea_b44, 
        itsea_b88, 
        itsea_c2, 
        itsea_e4
      ),
      na.rm = TRUE
    ) / n_itsea_agress,
    itsea_agress = if_else(
      n_itsea_agress < 10,
      NA_real_, itsea_agress
    ),
    itsea_depp = sum(
      c(
        itsea_a9, 
        itsea_a39, 
        itsea_a40, 
        itsea_b43, 
        itsea_b76, 
        itsea_b81, 
        itsea_b84, 
        itsea_b91, 
        itsea_b92
      ), na.rm = TRUE
    ) / n_itsea_depp,
    itsea_depp = if_else(
      n_itsea_depp < 7,
      NA_real_, itsea_depp
    ),
    itsea_empath = sum(
      c(
        itsea_a22, 
        itsea_b51, 
        itsea_b56, 
        itsea_b57, 
        itsea_b63, 
        itsea_b73, 
        itsea_c4
      ),
      na.rm = TRUE
    ) / n_itsea_empath,
    itsea_empath = if_else(
      n_itsea_empath < 6,
      NA_real_, itsea_empath
    ),
    itsea_gad = sum(
      c(
        itsea_a3, 
        itsea_a12, 
        itsea_a35, 
        itsea_b37, 
        itsea_b70, 
        itsea_b86, 
        itsea_e5, 
        itsea_e6, 
        itsea_e11
      ),
      na.rm = TRUE
    ) / n_itsea_gad,
    itsea_gad = if_else(
      n_itsea_gad < 8, 
      NA_real_, itsea_gad
    ),
    itsea_play = sum(
      c(
        itsea_a25, 
        itsea_a31, 
        itsea_b25, 
        itsea_b38, 
        itsea_b67, 
        itsea_b71
      ),
      na.rm = TRUE
    ) / n_itsea_play,
    itsea_play = if_else(
      n_itsea_play < 5,
      NA_real_, itsea_play
    ),
    itsea_negemo = sum(
      c(
        itsea_a7, 
        itsea_a21, 
        itsea_a23, 
        itsea_b31, 
        itsea_b45, 
        itsea_b50, 
        itsea_b53, 
        itsea_b59, 
        itsea_b65, 
        itsea_b66, 
        itsea_b74, 
        itsea_b80, 
        itsea_b85
      ),
      na.rm = TRUE
    ) / n_itsea_negemo,
    itsea_negemo = if_else(
      n_itsea_negemo < 10,
      NA_real_, itsea_negemo
    ),
    itsea_social = sum(
      c(
        itsea_a10, 
        itsea_a13, 
        itsea_a15, 
        itsea_b6, 
        itsea_b19, 
        itsea_b26, 
        itsea_b28, 
        itsea_b42, 
        itsea_b60, 
        itsea_b90
      ),
      na.rm = TRUE
    ) / n_itsea_social,
    itsea_social = if_else(
      n_itsea_social < 8, 
      NA_real_, itsea_social
    ),
    itsea_extern = sum(
      c(
      itsea_activity,
      itsea_agress
      ) / 2 
    ),
    itsea_extern = if_else(
      is.na(itsea_activity) | is.na(itsea_agress),
      NA_real_, itsea_extern
    ),
    itsea_intern = sum(
      c(
        itsea_depp,
        itsea_gad
      )
    ) / 2,
    itsea_intern = if_else(
      is.na(itsea_depp) | is.na(itsea_gad),
      NA_real_, itsea_intern
    )
  ) %>% 
  ungroup()
```

# Calculate symptom and competence scores
```{r}
itsea <-
  itsea %>%
  group_by(ID) %>% 
  mutate(
    itsea_symptoms = sum(
      c(
        itsea_intern,
        itsea_extern,
        itsea_negemo
        ), na.rm = TRUE
      ),
    itsea_competence = sum(
      c(
        itsea_play,
        itsea_empath,
        itsea_social
      ), na.rm = TRUE
    )
  ) %>% 
  ungroup() 
```

# Calculate alphas

## subscales
```{r}
# activity
itsea %>% 
  select(
    itsea_a2, 
    itsea_a4, 
    itsea_b8, 
    itsea_b10, 
    itsea_b27, 
    itsea_b32
  ) %>% 
  alpha()
```

```{r}
# aggression
itsea %>% 
  select(
    itsea_a28, 
    itsea_a30, 
    itsea_a33, 
    itsea_b3, 
    itsea_b9, 
    itsea_b16, 
    itsea_b30, 
    itsea_b34, 
    itsea_b44, 
    itsea_b88, 
    itsea_c2, 
    itsea_e4
  ) %>%   
  alpha()

```

```{r}
# depression
itsea %>% 
  select(
    itsea_a9, 
    itsea_a39, 
    itsea_a40, 
    itsea_b43, 
    itsea_b76, 
    itsea_b81, 
    itsea_b84, 
    itsea_b91, 
    itsea_b92
  ) %>% 
  alpha()
```

```{r}
# empathy
itsea %>% 
  select(
    itsea_a22, 
    itsea_b51, 
    itsea_b56, 
    itsea_b57, 
    itsea_b63, 
    itsea_b73, 
    itsea_c4
  ) %>% 
  alpha()
```

```{r}
# GAD
itsea %>% 
  select(
    itsea_a3, 
    itsea_a12, 
    itsea_a35, 
    itsea_b37, 
    itsea_b70, 
    itsea_b86, 
    itsea_e5, 
    itsea_e6, 
    itsea_e11
  ) %>% 
  alpha()
```

```{r}
# play
itsea %>% 
  select(
    itsea_a25, 
    itsea_a31, 
    itsea_b25, 
    itsea_b38, 
    itsea_b67, 
    itsea_b71
  ) %>% 
  alpha()
```

```{r}
# negative emotionality
itsea %>% 
  select(
    itsea_a7, 
    itsea_a21, 
    itsea_a23, 
    itsea_b31, 
    itsea_b45, 
    itsea_b50, 
    itsea_b53, 
    itsea_b59, 
    itsea_b65, 
    itsea_b66, 
    itsea_b74, 
    itsea_b80, 
    itsea_b85
  ) %>% 
  alpha()
```

```{r}
# pro-social
itsea %>% 
  select(
    itsea_a10, 
    itsea_a13, 
    itsea_a15, 
    itsea_b6, 
    itsea_b19, 
    itsea_b26, 
    itsea_b28, 
    itsea_b42, 
    itsea_b60, 
    itsea_b90
  ) %>% 
  alpha()
```

## internalizing and externalizing
```{r}
# internalizing
itsea %>% 
  select(
    itsea_a9, 
    itsea_a39, 
    itsea_a40, 
    itsea_b43, 
    itsea_b76, 
    itsea_b81, 
    itsea_b84, 
    itsea_b91, 
    itsea_b92,
    itsea_a3, 
    itsea_a12, 
    itsea_a35, 
    itsea_b37, 
    itsea_b70, 
    itsea_b86, 
    itsea_e5, 
    itsea_e6, 
    itsea_e11
  ) %>% 
  alpha()
```

```{r}
# externalizing
itsea %>% 
  select(
    itsea_a2, 
    itsea_a4, 
    itsea_b8, 
    itsea_b10, 
    itsea_b27, 
    itsea_b32,
    itsea_a28, 
    itsea_a30, 
    itsea_a33, 
    itsea_b3, 
    itsea_b9, 
    itsea_b16, 
    itsea_b30, 
    itsea_b34, 
    itsea_b44, 
    itsea_b88, 
    itsea_c2, 
    itsea_e4
  ) %>% 
  alpha()
```

## total symptoms
```{r}
#total symptoms 
itsea %>% 
  select(
    itsea_a2, 
    itsea_a4, 
    itsea_b8, 
    itsea_b10, 
    itsea_b27, 
    itsea_b32,
    itsea_a28, 
    itsea_a30, 
    itsea_a33, 
    itsea_b3, 
    itsea_b9, 
    itsea_b16, 
    itsea_b30, 
    itsea_b34, 
    itsea_b44, 
    itsea_b88, 
    #itsea_c2, # only show if mother endorses child combining words; mostly missing
    itsea_e4,
    itsea_a9, 
    itsea_a39, 
    itsea_a40, 
    itsea_b43, 
    itsea_b76, 
    itsea_b81, 
    itsea_b84, 
    itsea_b91, 
    itsea_b92,
    itsea_a3, 
    itsea_a12, 
    itsea_a35, 
    itsea_b37, 
    itsea_b70, 
    itsea_b86, 
    itsea_e5, 
    itsea_e6, 
    itsea_e11,
    itsea_a7, 
    itsea_a21, 
    itsea_a23, 
    itsea_b31, 
    itsea_b45, 
    itsea_b50, 
    itsea_b53, 
    itsea_b59, 
    itsea_b65, 
    itsea_b66, 
    itsea_b74, 
    itsea_b80, 
    itsea_b85
  ) %>% 
  alpha()
```

## total competence
```{r}
#total competence
itsea %>% 
  select(
    itsea_a25, 
    itsea_a31, 
    itsea_b25, 
    itsea_b38, 
    itsea_b67, 
    itsea_b71,
    itsea_a10, 
    itsea_a13, 
    itsea_a15, 
    itsea_b6, 
    itsea_b19, 
    itsea_b26, 
    itsea_b28, 
    itsea_b42, 
    itsea_b60, 
    itsea_b90,
    itsea_a22, 
    itsea_b51, 
    itsea_b56, 
    itsea_b57, 
    itsea_b63, 
    itsea_b73 
    #itsea_c4 # only show if mother endorses child combining words; mostly missing
  ) %>% 
  alpha()
```

## Calculate clinical cut scores
Cutoff scores drawn from ITSEA manual.
```{r}
itsea <-
  itsea %>% 
  group_by(ID) %>% 
  mutate(
    activity_concern = case_when(
      male == 1 & itsea_activity >= 1.42 ~ "concern",
      male == 0 & itsea_activity >= 1.28 ~ "concern",
      TRUE ~ "none"
    ),
    agress_concern = case_when(
      male == 1 & itsea_agress >= .87 ~ "concern",
      male == 0 & itsea_agress >= .81 ~ "concern",
      TRUE ~ "none"
    ),
    depp_concern = case_when(
      male == 1 & itsea_depp >= .36 ~ "concern",
      male == 0 & itsea_depp >= .34 ~ "concern",
      TRUE ~ "none"
    ),
    gad_concern = case_when(
      male == 1 & itsea_gad >= .55 ~ "concern",
      male == 0 & itsea_gad >= .44 ~ "concern",
      TRUE ~ "none"
    ),
    negemo_concern = case_when(
      male == 1 & itsea_negemo >= .91 ~ "concern",
      male == 0 & itsea_negemo >= 1.01 ~ "concern",
      TRUE ~ "none"
    ),
    play_concern = case_when(
      male == 1 & itsea_play <= .96 ~ "concern",
      male == 0 & itsea_play <= 1.07 ~ "concern",
      TRUE ~ "none"
    ),
    empathy_concern = case_when(
      male == 1 & itsea_empath <= .32 ~ "concern",
      male == 0 & itsea_empath <= .51 ~ "concern",
      TRUE ~ "none"
    ),
    social_concern = case_when(
      male == 1 & itsea_social <= 1.38 ~ "concern",
      male == 0 & itsea_social <= 1.35 ~ "concern",
      TRUE ~ "none"
    ),
    itsea_concern = if_else(
      activity_concern == "concern" | agress_concern == "concern" |
        depp_concern == "concern" | gad_concern == "concern" |
        negemo_concern == "concern" | play_concern == "concern" |
        empathy_concern == "concern" | social_concern == "concern",
      "concern", "none"
    ),
    itsea_concern_ied = if_else(
      activity_concern == "concern" | agress_concern == "concern" |
        depp_concern == "concern" | gad_concern == "concern" |
        negemo_concern == "concern",
      "concern", "none"
    )
  ) %>% 
  ungroup()
```

#Calculate ITSEA concern counts
```{r}
q_18month_scored %>% 
  count(activity_concern)

q_18month_scored %>% 
  count(agress_concern)

q_18month_scored %>% 
  count(depp_concern)

q_18month_scored %>% 
  count(gad_concern)

q_18month_scored %>% 
  count(negemo_concern)

q_18month_scored %>% 
  count(empathy_concern)

q_18month_scored %>% 
  count(social_concern)

q_18month_scored %>% 
  count(play_concern)

q_18month_scored %>% 
  count(itsea_concern) %>% 
  mutate(prop = n / sum(n))

q_18month_scored %>% 
  count(itsea_concern_ied) %>% 
  mutate(prop = n / sum(n))
```

# Visualizing subscale scores
```{r}
itsea %>% 
  select(itsea_activity:itsea_social) %>% 
  gather(itsea_subscale, score, itsea_activity:itsea_social) %>% 
  ggplot(aes(score, color = itsea_subscale)) +
  geom_freqpoly(bins = 5, size = 2, alpha = 1/2) +
  theme_minimal()

itsea %>% 
  select(
    Activity = itsea_activity,
    Aggression = itsea_agress,
    Depression = itsea_depp,
    Anxiety = itsea_gad,
    `Negative emotion` = itsea_negemo
  ) %>% 
  gather(itsea_subscale, score, Activity:`Negative emotion`) %>% 
  ggplot(aes(score)) +
  geom_histogram(bins = 12) +
  theme_lena +
  theme(
    strip.text = element_text(size = 16)
  ) +
  facet_wrap(.~itsea_subscale, scales = "free") +
  labs(
    x = "Subscale score"
  )

ggsave(
  "~/Desktop/BABIES/lena_symptoms/manuscript/itsea_subscale_hist.png",
  width = 7,
  height = 5
)


itsea %>% 
  select(itsea_activity:itsea_social) %>% 
  gather(itsea_subscale, score, itsea_activity:itsea_social) %>% 
  ggplot(aes(score)) +
  geom_histogram(bins = 12) +
  theme_minimal() +
  facet_wrap(.~itsea_subscale)

```

# ITSEA EFA
```{r}
itsea_subscales <- 
  itsea %>% 
  arrange(ID) %>% 
  select( 
    itsea_activity:itsea_social
    ) %>% 
  na.omit()

itsea_sub_corr <-
  cor(itsea_subscales)
```

```{r}
fa.parallel(itsea_subscales)

itsea_fa2 <- fa(
  itsea_subscales, 
  nfactors = 2, 
  rotate = "varimax"
  )
itsea_fa2
itsea_fa2$Vaccounted
```

#### Heatmap of 2-factor solution

```{r}
factor_names = c(
  "Factor 1:\nSymptoms", 
  "Factor 2:\nCompetence"
  )

loadings <-
  itsea_fa2$loadings[] %>% 
  data.frame() %>% 
  rownames_to_column("measure") %>% 
  gather(factor, loading, -measure) %>%  
  mutate(factor = as.character(factor(factor, labels = factor_names))) 
  
# get fa.sort() order
order <- 
  loadings %>%
  group_by(measure) %>%
  top_n(1, abs(loading)) %>%
  ungroup() %>%
  arrange(desc(factor), abs(loading)) %>%
  mutate(order = 1:length(levels(factor(loadings$measure)))) %>%
  select(measure, order)
  
# get percent shared variance explained
shared_var <- itsea_fa2$Vaccounted %>%
  data.frame() %>%
  rownames_to_column("stat") %>%
  filter(stat == "Proportion Var") %>%
  select(-stat) %>%
  gather(factor, var) %>%
  mutate(factor = as.character(factor(factor, labels = factor_names))) %>%
  mutate(var = paste0(factor, "\n(", round(var, 2)*100, "% var.)"))

  # make plot
loadings_plot_data <- 
  loadings %>% 
  left_join(order, by = "measure") %>%
  left_join(shared_var, by = "factor")

loadings_plot_data %>% 
  mutate(
    measure = fct_recode(
      measure,
      Activity = "itsea_activity", 
      Aggression = "itsea_agress", 
      Depression = "itsea_depp", 
      Anxiety = "itsea_gad", 
      Play = "itsea_play",
      Empathy = "itsea_empath",
      `Negative emotionality` = "itsea_negemo",
      `Pro-social relations` = "itsea_social"
    )
  ) %>% 
  ggplot(
    aes(
      x = var, 
      y = reorder(measure, order), 
      fill = loading,
      label = round(loading, 2)
    )
  ) +
  geom_tile(color = "black") +
  geom_text(size = 6, color = "black") +
  scale_fill_distiller(
    limits = c(-1, 1), 
    palette = "RdYlBu",
    guide = guide_colorbar(barheight = 20)
  ) +
  scale_x_discrete(position = "top") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    axis.title = element_blank(), 
    axis.text.y = element_text(size = 18),
    axis.text.x = element_text(size = 17)
  )

ggsave("~/Desktop/BABIES/lena_symptoms/manuscript/2efa.png", height = 10, width = 11.5)
```

## Export data

```{r}
q_18month_scored <-
  itsea %>% 
  mutate(
    survey_date = parse_date_time(survey_date, orders = c("ymd", "mdy")),
    baby_dob = parse_date_time(baby_dob, orders = c("ymd", "mdy")),
    baby_age_18month = (baby_dob %--% survey_date) / months(1),
    baby_age_18month_int = as.integer(baby_age_18month)
  ) 

glimpse(q_18month_scored)

write_csv(q_18month_scored, "~/Desktop/BABIES/lena_symptoms/data_final/q_18month_scores_20190726.csv")
```


