---
title: "Score 18-month questionnaires"
output: html_notebook
---

```{r}
# Libraries
library(tidyverse)
library(lubridate)
library(dplyr)
library(purrr)
#install.packages("psych")
library(psych)
#install.packages("GPArotation")
library(GPArotation)

# Files

## Fran file paths
q_18month_cs_file <- "~/Desktop/lena_18mo_symptoms/Data/qs_18mo_cs_datecleaned_20190726.csv"
q_18month_long_file <- "~/Desktop/lena_18mo_symptoms/Data/qs_18mo_long_datecleaned_20190726.csv"
demographics_file <- "~/Desktop/lena_18mo_symptoms/Data/demographics_20190726.csv"

## Lucy file paths
# q_18month_cs_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/qs_18mo_cs_datecleaned_20190510.csv"
# q_18month_long_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/qs_18mo_long_datecleaned_20190510.csv"
# demographics_file <- "~/Desktop/BABIES/lena_symptoms/fran_box_data/demographics_20190510.csv"
```

**Read in data**
```{r}
q_18month <-
  read_csv(q_18month_cs_file) %>% 
  rename(ID = record_id) %>% 
  mutate(ID = as.integer(ID)) %>% 
  select(-redcap_event_name:-questionnaires_18mo_timestamp) %>% 
  bind_rows(
    read_csv(q_18month_long_file) %>% 
      rename(ID = record_id) %>% 
      mutate(ID = as.integer(ID)) %>%
      select(-redcap_event_name:-month_survey_timestamp) 
  ) %>% 
  left_join(
    read_csv(demographics_file) %>% 
      select(ID, male, baby_dob),
    by = "ID"
  )
```

# Score ITSEA

## Compute number of non-missing items for each subscale
```{r}
q_18month_scored <-
  q_18month %>% 
  group_by(ID) %>% 
  mutate(
    n_itsea_activity = sum(
      !is.na(
        c(
          itsea_a2, 
          itsea_a4, 
          itsea_b8, 
          itsea_b10, 
          itsea_b27, 
          itsea_b32
        )
      )
    ),
    n_itsea_agress = sum(
      !is.na(
        c(
          itsea_a28, 
          itsea_a30, 
          itsea_a33, 
          itsea_b3, 
          itsea_b9, 
          itsea_b16, 
          itsea_b30, 
          itsea_b34, 
          itsea_b44, 
          itsea_b88, 
          itsea_c2, 
          itsea_e4
        )
      )
    ),
    n_itsea_depp = sum(
      !is.na(
        c(
          itsea_a9, 
          itsea_a39, 
          itsea_a40, 
          itsea_b43, 
          itsea_b76, 
          itsea_b81, 
          itsea_b84, 
          itsea_b91, 
          itsea_b92
        )
      )
    ),
    n_itsea_empath = sum(
      !is.na(
        c(
          itsea_a22, 
          itsea_b51, 
          itsea_b56, 
          itsea_b57, 
          itsea_b63, 
          itsea_b73, 
          itsea_c4
        )
      )
    ),
    n_itsea_gad = sum(
      !is.na(
        c(
          itsea_a3, 
          itsea_a12, 
          itsea_a35, 
          itsea_b37, 
          itsea_b70, 
          itsea_b86, 
          itsea_e5, 
          itsea_e6, 
          itsea_e11
        )
      )
    ),
    n_itsea_play = sum(
      !is.na(
        c(
          itsea_a25, 
          itsea_a31, 
          itsea_b25, 
          itsea_b38, 
          itsea_b67, 
          itsea_b71
        )
      )
    ),
    n_itsea_negemo = sum(
      !is.na(
        c(
          itsea_a7, 
          itsea_a21, 
          itsea_a23, 
          itsea_b31, 
          itsea_b45, 
          itsea_b50, 
          itsea_b53, 
          itsea_b59, 
          itsea_b65, 
          itsea_b66, 
          itsea_b74, 
          itsea_b80, 
          itsea_b85
        )
      )
    ),
    n_itsea_social = sum(
      !is.na(
        c(
          itsea_a10, 
          itsea_a13, 
          itsea_a15, 
          itsea_b6, 
          itsea_b19, 
          itsea_b26, 
          itsea_b28, 
          itsea_b42, 
          itsea_b60, 
          itsea_b90
        )
      )
    )
  ) %>% 
  ungroup()
```

## Calculate subscale scores
```{r}
q_18month_scored <-
  q_18month_scored %>% 
  #reverse code items
  mutate_at(
    vars(
      itsea_b43, itsea_b65
    ),
    funs(
      . %>%
        dplyr::recode(
          "0" = 2, 
          "1" = 1, 
          "2" = 0
        )
    )
  ) %>% 
  group_by(ID) %>% 
  mutate(
    itsea_activity = sum(
      c(
        itsea_a2, 
        itsea_a4, 
        itsea_b8, 
        itsea_b10, 
        itsea_b27, 
        itsea_b32
      ), na.rm = TRUE
    ) / n_itsea_activity,
    itsea_activity = if_else(
      n_itsea_activity < 5, 
      NA_real_, itsea_activity
    ),
    itsea_agress = sum(
      c(
        itsea_a28, 
        itsea_a30, 
        itsea_a33, 
        itsea_b3, 
        itsea_b9, 
        itsea_b16, 
        itsea_b30, 
        itsea_b34, 
        itsea_b44, 
        itsea_b88, 
        itsea_c2, 
        itsea_e4
      ),
      na.rm = TRUE
    ) / n_itsea_agress,
    itsea_agress = if_else(
      n_itsea_agress < 10,
      NA_real_, itsea_agress
    ),
    itsea_depp = sum(
      c(
        itsea_a9, 
        itsea_a39, 
        itsea_a40, 
        itsea_b43, 
        itsea_b76, 
        itsea_b81, 
        itsea_b84, 
        itsea_b91, 
        itsea_b92
      ), na.rm = TRUE
    ) / n_itsea_depp,
    itsea_depp = if_else(
      n_itsea_depp < 7,
      NA_real_, itsea_depp
    ),
    itsea_empath = sum(
      c(
        itsea_a22, 
        itsea_b51, 
        itsea_b56, 
        itsea_b57, 
        itsea_b63, 
        itsea_b73, 
        itsea_c4
      ),
      na.rm = TRUE
    ) / n_itsea_empath,
    itsea_empath = if_else(
      n_itsea_empath < 6,
      NA_real_, itsea_empath
    ),
    itsea_gad = sum(
      c(
        itsea_a3, 
        itsea_a12, 
        itsea_a35, 
        itsea_b37, 
        itsea_b70, 
        itsea_b86, 
        itsea_e5, 
        itsea_e6, 
        itsea_e11
      ),
      na.rm = TRUE
    ) / n_itsea_gad,
    itsea_gad = if_else(
      n_itsea_gad < 8, 
      NA_real_, itsea_gad
    ),
    itsea_play = sum(
      c(
        itsea_a25, 
        itsea_a31, 
        itsea_b25, 
        itsea_b38, 
        itsea_b67, 
        itsea_b71
      ),
      na.rm = TRUE
    ) / n_itsea_play,
    itsea_play = if_else(
      n_itsea_play < 5,
      NA_real_, itsea_play
    ),
    itsea_negemo = sum(
      c(
        itsea_a7, 
        itsea_a21, 
        itsea_a23, 
        itsea_b31, 
        itsea_b45, 
        itsea_b50, 
        itsea_b53, 
        itsea_b59, 
        itsea_b65, 
        itsea_b66, 
        itsea_b74, 
        itsea_b80, 
        itsea_b85
      ),
      na.rm = TRUE
    ) / n_itsea_negemo,
    itsea_negemo = if_else(
      n_itsea_negemo < 10,
      NA_real_, itsea_negemo
    ),
    itsea_social = sum(
      c(
        itsea_a10, 
        itsea_a13, 
        itsea_a15, 
        itsea_b6, 
        itsea_b19, 
        itsea_b26, 
        itsea_b28, 
        itsea_b42, 
        itsea_b60, 
        itsea_b90
      ),
      na.rm = TRUE
    ) / n_itsea_social,
    itsea_social = if_else(
      n_itsea_social < 8, 
      NA_real_, itsea_social
    ),
    itsea_extern = sum(
      c(
      itsea_activity,
      itsea_agress
      ) / 2 
    ),
    itsea_extern = if_else(
      is.na(itsea_activity) | is.na(itsea_agress),
      NA_real_, itsea_extern
    ),
    itsea_intern = sum(
      c(
        itsea_depp,
        itsea_gad
      )
    ) / 2,
    itsea_intern = if_else(
      is.na(itsea_depp) | is.na(itsea_gad),
      NA_real_, itsea_intern
    )
  ) %>% 
  ungroup()
```

# Calculate symptom and competence scores
```{r}
q_18month_scored <-
  q_18month_scored %>%
  group_by(ID) %>% 
  mutate(
    itsea_symptoms = sum(
      c(
        itsea_intern,
        itsea_extern,
        itsea_negemo
        ), na.rm = TRUE
      ),
    itsea_competence = sum(
      c(
        itsea_play,
        itsea_empath,
        itsea_social
      ), na.rm = TRUE
    )
  ) %>% 
  ungroup() 
```

#Calculate alpha
```{r}
itsea_alpha <-
q_18month_scored %>% 
  select(
    itsea_intern,
    itsea_extern,
    itsea_negemo
  )

psych::alpha(itsea_alpha)

```

## Calculate clinical cut scores
```{r}
q_18month_scored <-
  q_18month_scored %>% 
  group_by(ID) %>% 
  mutate(
    activity_concern = case_when(
      male == 1 & itsea_activity >= 1.42 ~ "concern",
      male == 0 & itsea_activity >= 1.28 ~ "concern",
      TRUE ~ "none"
    ),
    agress_concern = case_when(
      male == 1 & itsea_agress >= .87 ~ "concern",
      male == 0 & itsea_agress >= .81 ~ "concern",
      TRUE ~ "none"
    ),
    depp_concern = case_when(
      male == 1 & itsea_depp >= .36 ~ "concern",
      male == 0 & itsea_depp >= .34 ~ "concern",
      TRUE ~ "none"
    ),
    gad_concern = case_when(
      male == 1 & itsea_gad >= .55 ~ "concern",
      male == 0 & itsea_gad >= .44 ~ "concern",
      TRUE ~ "none"
    ),
    negemo_concern = case_when(
      male == 1 & itsea_negemo >= .91 ~ "concern",
      male == 0 & itsea_negemo >= 1.01 ~ "concern",
      TRUE ~ "none"
    ),
    play_concern = case_when(
      male == 1 & itsea_play <= .96 ~ "concern",
      male == 0 & itsea_play <= 1.07 ~ "concern",
      TRUE ~ "none"
    ),
    empathy_concern = case_when(
      male == 1 & itsea_empath <= .32 ~ "concern",
      male == 0 & itsea_empath <= .51 ~ "concern",
      TRUE ~ "none"
    ),
    social_concern = case_when(
      male == 1 & itsea_social <= 1.38 ~ "concern",
      male == 0 & itsea_social <= 1.35 ~ "concern",
      TRUE ~ "none"
    ),
    itsea_concern = if_else(
      activity_concern == "concern" | agress_concern == "concern" |
        depp_concern == "concern" | gad_concern == "concern" |
        negemo_concern == "concern" | play_concern == "concern" |
        empathy_concern == "concern" | social_concern == "concern",
      "concern", "none"
    ),
    itsea_concern_ied = if_else(
      activity_concern == "concern" | agress_concern == "concern" |
        depp_concern == "concern" | gad_concern == "concern" |
        negemo_concern == "concern",
      "concern", "none"
    )
  ) %>% 
  ungroup()
```

#Calculate ITSEA concern counts
```{r}
q_18month_scored %>% 
  count(activity_concern)

q_18month_scored %>% 
  count(agress_concern)

q_18month_scored %>% 
  count(depp_concern)

q_18month_scored %>% 
  count(gad_concern)

q_18month_scored %>% 
  count(negemo_concern)

q_18month_scored %>% 
  count(empathy_concern)

q_18month_scored %>% 
  count(social_concern)

q_18month_scored %>% 
  count(play_concern)

q_18month_scored %>% 
  count(itsea_concern) %>% 
  mutate(prop = n / sum(n))

q_18month_scored %>% 
  count(itsea_concern_ied) %>% 
  mutate(prop = n / sum(n))
```

# Visualizing subscale scores
```{r}
q_18month_scored %>% 
  select(itsea_activity:itsea_social) %>% 
  gather(itsea_subscale, score, itsea_activity:itsea_social) %>% 
  ggplot(aes(score, color = itsea_subscale)) +
  geom_freqpoly(bins = 5, size = 2, alpha = 1/2) +
  theme_minimal()

q_18month_scored %>% 
  select(itsea_activity:itsea_social) %>% 
  gather(itsea_subscale, score, itsea_activity:itsea_social) %>% 
  ggplot(aes(score)) +
  geom_histogram(bins = 12) +
  theme_minimal() +
  facet_wrap(.~itsea_subscale)

```



# ITSEA EFA
```{r}
itsea_subscales <- 
  q_18month_scored %>% 
  arrange(ID) %>% 
  select( 
    itsea_activity:itsea_social
    ) %>% 
  na.omit()

itsea_sub_corr <-
  cor(itsea_subscales)

VSS.scree(itsea_subscales)

itsea_fa2 <- fa(
  itsea_sub_corr, 
  nfactors = 2, 
  rotate = "oblimin",
  scores = "tenBerge",
  impute = "median"
  )
itsea_fa2
itsea_fa2$Vaccounted

itsea_fa3 <- fa(
  itsea_sub_corr, 
  nfactors = 3, 
  rotate = "oblimin", #allows factors to be correlated
  scores = "tenBerge", #allows scores to be correlated
  impute = "median"
  )
itsea_fa3
itsea_fa3$Vaccounted
```
Retaining 2 factors given eigen value given third factor explained small amount of variance and eigen value < 1. 

Factor 1 explains 24% of the variance and is basically "symptoms."
Factor 2 explains 20% of the variance and is basically "competence."

## Extract factor scores
```{r}
itsea_fascores_calc <-
  factor.scores(
    itsea_subscales, 
    itsea_fa2, 
    method = "tenBerge", 
    impute = "median"
  ) 

itsea_fascores_pull <-
  itsea_fascores_calc$scores %>% 
  as_tibble() %>% 
  rownames_to_column("index")

itsea_fascores <-
  q_18month_scored %>% 
  arrange(ID) %>% 
  select(
    ID,
    itsea_activity:itsea_social
  ) %>% 
  rownames_to_column("index")

itsea_fascores <-
  itsea_fascores %>% 
  left_join(itsea_fascores_pull, by = "index") %>% 
  select(-index) %>% 
  rename(
    itsea_f1_symptoms = MR1,
    itsea_f2_competence = MR2
  )
```

## Export data

```{r}
q_18month_scored <-
  q_18month_scored %>% 
  mutate(
    survey_date = parse_date_time(survey_date, orders = c("ymd", "mdy")),
    baby_age_18month = (baby_dob %--% survey_date) / months(1),
    baby_age_18month_int = as.integer(baby_age_18month)
  ) 


write_csv(q_18month_scored, "~/Desktop/lena_18mo_symptoms/Data_Final/q_18month_scores_20190726.csv")
```


